### 2018-11-19 Vortrag/slides/sections/ddplus.tex
First version of motivation completed.

\subsection{The dd+ algorithm}
\begin{frame}{The dd+ algorithm}
	\begin{center}
		Solves it all!
	\end{center}
\end{frame}

### 2018-11-19 Vortrag/slides/sections/difficulties.tex
First version of motivation completed.

\subsection{Interference}
\begin{frame}{Difficulty \hashtag 1: Interference}
	\begin{itemize}
		\item $ddsimple$ works for single failure inducing-changes.
			\begin{itemize}
				\item In each recursion step it applies only the set of changes known to contain a failure inducing change.
			\end{itemize}
		\item But what if two changes exist that individually pass the tests but their combination induces failure?
	\end{itemize}

	\begin{alertblock}{Difficulty \hashtag 1: Interference}
		Let $c_1,c_2 \in \C$. $c_1$ and $c_2$ \textbf{interfere} when $test(c_1)=\cmark$, $test(c_2)=\cmark$ but $test(c_1\cup c_2) = \xmark$.
	\end{alertblock}

	\begin{exampleblock}{Idea: Leave one set of changes applied}
		If a configuration $c = c_1 \cup c_2$ with $test(c_1)=\cmark$ and $test(c_2)=\cmark$ is found by $simpledd$ then run $simpledd$ on $c_1$ while leaving $c_2$ applied and vice versa.
	\end{exampleblock}
\end{frame}

\subsection{Inconsistency}
\begin{frame}{Difficulty \hashtag 2: Inconsistency}

\end{frame}

### 2018-11-19 Vortrag/slides/sections/difficulties.tex
Completed first version of difficulties.

@@
 
+\end{frame}
+
+\begin{frame}{Difficulty \hashtag 1: Interference}
 	\begin{exampleblock}{Idea: Leave one set of changes applied}
-		If a configuration $c = c_1 \cup c_2$ with $test(c_1)=\cmark$ and $test(c_2)=\cmark$ is found by $simpledd$ then run $simpledd$ on $c_1$ while leaving $c_2$ applied and vice versa.
+		If a configuration $c = c_1 \cup c_2$ with $test(c_1)=\cmark$ and $test(c_2)=\cmark$ is found by $simpledd$ interference between $c_1$ and $c_2$ has been detected.
+Run $simpledd$ on $c_1$ while leaving $c_2$ applied and vice versa.
 	\end{exampleblock}
+
+	\bigskip
+
+	%If we add this case to the $simpledd$ algorithm it leaves us with Andreas Zellers $dd$ algorithm: $dd(c: 2^{\C}) = dd_2(c,\emptyset) \text{, where}$\\[1em]
+
+	\begin{algorithmic}[1]
+		\Function{$dd_2$}{$c,r:2^\C$} $: 2^{\C}$
+		\State let $c_1,c_2 \subseteq c \text{ with } c_1 \cup c_2 = c, c_1 \cap c_2 = \emptyset, |c_1| \approx |c_2|$
+		\State \Return $ \begin{cases}
+				c & \text{if } |c|=1, \\
+				dd_2(c_1,r) & \text{if } test(c_1 \cup r) = \xmark,\\
+				dd_2(c_2,r) & \text{if } test(c_2 \cup r) = \xmark, \\
+				dd_2(c_1, c_2 \cup r) \cup dd_2(c_2, c_1 \cup r) & \text{otherwise} 
+			\end{cases}$
+		\EndFunction
+	\end{algorithmic}
+
 \end{frame}
@@
 \begin{frame}{Difficulty \hashtag 2: Inconsistency}
+	\begin{itemize}
+		\item $dd$ combines changes arbitrarily (in the case of interference)
+		\item That can lead to inconsistent configurations, i.e. no test outcome can be determined for these configurations 
+	\end{itemize}
+
+	\begin{alertblock}{Difficulty \hashtag 2: Inconsistency}
+		Let $c_1,c_2 \subseteq \C$. An \textbf{inconsistency} occurs when $test(c_1 \cup c_2) = \qmark$.
+	\end{alertblock}
+	\bigskip
+	\begin{exampleblock}{Idea: More granular subsets}
+		If less changes are applied at once the chances of an inconsistent result are reduced. Hence, if the algorithm cannot find any consistent confgurations reduce the number of changes per subset.
+	\end{exampleblock}
 
 \end{frame}
+
+\begin{frame}{Difficulty \hashtag 2: Inconsistency}
+
+	Necessary changes to $dd$:
+	\begin{enumerate}
+		\item Extend $dd$ to work on a number $n$ of subsets $c_1, \dots, c_n$
+		\item \textbf{Interference} occurs when $c_i$ and its complement $\bar{c_i}$ both pass: $test(c_i) = \cmark$ and $test(\bar{c_i}) = \cmark$ ($\bar{c_i}=\C \setminus c_i$)
+		\item Add the case of \textbf{preference}: If $test(c_i)=\qmark$ and $test(\bar{c_i})=\cmark$ we deduce that $c_i$ contains a failure inducing change. 
+		\item Add the case of \textbf{Try again}: In any other case repeat the process with $2n$ subsets.
+	\end{enumerate}
+\end{frame}

### 2018-11-27 Vortrag/slides/sections/change-effect-chains.tex
Finished first version of slides. Included Change-Effect-Chain-Isolation.

\subsection{The DD principle as a tool}
\begin{frame}{Using DD to find the reason a program fails}
	\begin{itemize}
		\item The \ddp\ algorithm as described above can be used to find out which part of a certain input is responsible for crashing a program.
		\item From a developers perspective it would be great to know \textit{why} this specific input induces failure.
	\end{itemize}

\end{frame}

\begin{frame}{Using DD to find the reason a program fails}
	\begin{alertblock}{Difficulty: Causalities from the input change to the failure}
		A minimal change in the input has major consequences for the program execution. How can the debugger know which chain of effects is responsible for the eventual failure?
	\end{alertblock}

	\pause

	\begin{exampleblock}{Idea: Delta Debugging on program states}
		Use delta debugging on program states of a known succeeding run $r_\cmark$ and a known failing run $r_\xmark$ of the program. For each analyzed pair of program states we would then know what current difference causes the failure down the line.

		\bigskip

		$\Rightarrow$ A \textit{Cause-Effect-Chain} can be constructed from those comparisons.
	\end{exampleblock}	
\end{frame}

\begin{frame}{Delta Debugging on program states}
	\begin{itemize}
		\item A program state is essentially a set of $(variable, value)$ pairs.
		\item A program state of $r_\cmark$ and one of $r_\xmark$ can have the following differences (the deltas): \begin{itemize}
			\item A variable only present in one state
			\item A different value of a variable present in both states
		\end{itemize}
		\item With this the runs $r_\cmark$ and $r_\xmark$ can be compared in different significant locations. 
		\item We use delta debugging to find the current $(variable, value)$ pairs that are responsible for the failure of $r_\xmark$ further down the line.
	\end{itemize}

	%\bigskip

	%$\Rightarrow$ This allows us to build a \textit{Cause-Effect-Chain} that helps understand why the program fails.
\end{frame}

\newcommand{\cecarrow}{\to}

\begin{frame}{\textit{HOWCOME} on $fail.c$}

	Zellers prototypical \textit{HOWCOME} algorithm returns the following Cause-Effect-Chain for $fail.c$:
	\small
	\begin{enumerate}
		\item Execution reaches \textbf{main}. \\Since the program was invoked as ``ccl -O fail.i'' variable \textbf{$argv[2]$} is now \textbf{``fail.i''}
		\item Execution reaches \textbf{combine\_instructions}. \\ Since $argv[2]$ was ``fail.i'', variable \textbf{$*first\_loop\_store\_insn\cecarrow fld[1].rtx \cecarrow fld[1].rtx \cecarrow fld[3].rtx \cecarrow fld[1].rtx$} is now \textbf{$\langle new rtx\_def\rangle$}.
		\item Execution reaches \textbf{if\_then\_else\_cond (95th hit)}. \\ Since $*first\_loop\_store\_insn\cecarrow fld[1].rtx \cecarrow fld[1].rtx \cecarrow fld[3].rtx \cecarrow fld[1].rtx$ was $\langle new rtx\_def\rangle$, variable \textbf{$link \cecarrow fld[0].rtx \cecarrow fld[0].rtx$} is now \textbf{$link$}.
		\item Execution ends. \\ Since variable $link \cecarrow fld[0].rtx \cecarrow fld[0].rtx$ was $link$, the program now \textbf{terminates with a SIGSEGV signal}. The program fails.
	\end{enumerate}

\end{frame}

### 2018-11-27 Vortrag/slides/sections/ddplus.tex
Finished first version of slides. Included Change-Effect-Chain-Isolation.

@@
 \subsection{The dd+ algorithm}
-\begin{frame}{The dd+ algorithm}
-	\begin{center}
-		Solves it all!
-	\end{center}
+
+\newcommand{\ddp}{$dd^{+}$}
+
+\begin{frame}{The \ddp\ algorithm}
+	Zellers \textbf{extended} $dd$ algorithm \ddp\ deals with the following cases:
+	\begin{enumerate}
+		\item ``found (in $c_i$)'' \only<2->{$\Rightarrow$ continue search in $c_i$}
+		\item ``interference'' \only<3->{$\Rightarrow$ search in $c_i$ while leaving $\bar{c_i}$ applied and vice versa}
+		\item ``preference'' \only<4->{$\Rightarrow$ search in $c_i$ while leaving $\bar{c_i}$ applied}
+		\item ``try again'' \only<5->{$\Rightarrow$ repeat search on same change set with twice as many subsets}
+		%\item ``nothing left		
+	\end{enumerate}
 \end{frame}
+
+\begin{frame}{The \ddp\ algorithm}
+	Properties of \ddp:
+	\begin{itemize}
+		\item \ddp finds a minimal failure-inducing subset of changes as long as the changes are safe, i.e. combined they result in a consistent configuration.
+		\item \ddp has at most linear time complexity like $dd$.
+	\end{itemize}
+\end{frame}
+
+\begin{frame}{Use case for \ddp\: Analyzing input that crashes a program}
+	Consider the C program $fail.c$ that crashed GCC in version 2.95.2:
+
+	\begin{columns}
+		\begin{column}{0.475\textwidth}
+			\begin{align*}
+	     &\text{double \textbf{mult}(double z[], int n)}\\
+	     &\{\\
+	     	&\quad \text{int } i,j; \\
+	     	&\quad i = 0; \\
+	     	&\quad for(j = 0; j < n; j++)\{ \\
+	     	&\quad 	\quad i = i + j + 1; \\
+	     &	\quad 	\quad z[i] = z[i] * (z[0] + 1.0); \\
+	     &	\quad \} \\
+	     &	\quad \text{return } z[n]; \\
+	    & \}
+	\end{align*} 
+		\end{column}
+		\begin{column}{0.475\textwidth}
+
+		\end{column}
+
+	\end{columns}
+
+\end{frame}
+
+\begin{frame}{How \ddp\ finds the minimal set of C tokens that causes the crash}
+	\begin{tabular}{c l c}
+				\hashtag & GCC input & test \\
+				\hline
+				1 & double mult($\dots$) $\{ int \; i, j; \; i = 0; \; for (\dots) \{ \dots\} \dots\}$ & $\xmark$ \\
+				2 &\gray{ double mult($\dots$) $\{ int \; i, j; \; i = 0; \; for (\dots) \{ \dots\} \dots\}$} & $\cmark$ \\
+				3 & double mult($\dots$) $\{\gray{ int \; i, j; \; i = 0; \; for (\dots) \{ \dots\} \dots}\}$ & $\cmark$ \\
+				4 & double mult($\dots$) $\{ int \; i, j; \; i = 0; \; \gray{for (\dots) \{ \dots\} \dots}\}$ & $\cmark$ \\
+				5 & double mult($\dots$) $\{ int \; i, j; \; i = 0; \; for (\dots) \{ \dots\}\gray{ \dots}\}$ & $\xmark$ \\
+				6 & double mult($\dots$) $\{ int \; i, j; \; i = 0; \; for (\gray{\dots}) \{ \dots\} \gray{\dots}\}$ & $\cmark$ \\
+
+				\vdots & \multicolumn{1}{c}{\vdots} & \vdots \\
+				18 & \multicolumn{1}{c}{\dots \quad $z[i] = z[i] * (z[0] + 1.0);$ \quad \dots } & \xmark \\
+				19 & \multicolumn{1}{c}{\dots \quad $z[i] = z[i] * (z[0] \gray{+ 1.0});$ \quad \dots } & \cmark \\
+				20 & \multicolumn{1}{c}{\dots \quad $z[i] = z[i] * (z[0] \gray{+} 1.0);$ \quad \dots } & \qmark \\
+				21 & \multicolumn{1}{c}{\dots \quad $z[i] = z[i] * (z[0] + \gray{1.0});$ \quad \dots } & \qmark \\
+		\end{tabular}
+
+\end{frame}

### 2018-11-27 Vortrag/slides/sections/difficulties.tex
Finished first version of slides. Included Change-Effect-Chain-Isolation.

@@
 		\item $dd$ combines changes arbitrarily (in the case of interference)
-		\item That can lead to inconsistent configurations, i.e. no test outcome can be determined for these configurations 
+		\item This can lead to inconsistent configurations, i.e. no test outcome can be determined for these configurations 
 	\end{itemize}
@@
 		\item Add the case of \textbf{preference}: If $test(c_i)=\qmark$ and $test(\bar{c_i})=\cmark$ we deduce that $c_i$ contains a failure inducing change. 
-		\item Add the case of \textbf{Try again}: In any other case repeat the process with $2n$ subsets.
+		\item Add the case of \textbf{Try again}: In any other case repeat the process with $2n$ subsets to improve the chance for consistent configurations.
 	\end{enumerate}

### 2018-12-14 Vortrag/slides/sections/ddplus.tex
Made end of slides nicer.

@@
 \begin{frame}{The \ddp\ algorithm}
-	Zellers \textbf{extended} $dd$ algorithm \ddp\ deals with the following cases:
+	Zeller's \textbf{extended} $dd$ algorithm \ddp\ deals with the following cases:
 	\begin{enumerate}
@@
 		\begin{column}{0.475\textwidth}
+			\texttt{
 			\begin{align*}
-	     &\text{double \textbf{mult}(double z[], int n)}\\
+	     &\text{double \textbf{mult}(double $z[]$, int $n$)}\\
 	     &\{\\
@@
 	     	&\quad i = 0; \\
-	     	&\quad for(j = 0; j < n; j++)\{ \\
+	     	&\quad \text{for}(j = 0; j < n; j++)\{ \\
 	     	&\quad 	\quad i = i + j + 1; \\
@@
 	\end{align*}
+		} 
 		\end{column}

### 2018-12-14 Vortrag/slides/sections/cause-effect-chains.tex

--- 2018-11-27 Vortrag/slides/sections/change-effect-chains.tex
@@
 
-\begin{frame}{\textit{HOWCOME} on $fail.c$}
+\begin{frame}{\textit{HOWCOME} on \texttt{fail.c}}
 
-	Zellers prototypical \textit{HOWCOME} algorithm returns the following Cause-Effect-Chain for $fail.c$:
-	\small
-	\begin{enumerate}
+	\begin{columns}
+		\begin{column}{0.4\textwidth}
+
+		\begin{overpic}[width=0.95\textwidth]{../figures/check-ripped.png}
+		\put (0,140) { \parbox{0.8\textwidth}{
+			\tiny
+		\begin{itemize}
 			\item Execution reaches \textbf{main}. \\Since the program was invoked as ``ccl -O fail.i'' variable \textbf{$argv[2]$} is now \textbf{``fail.i''}
@@
 			\item Execution ends. \\ Since variable $link \cecarrow fld[0].rtx \cecarrow fld[0].rtx$ was $link$, the program now \textbf{terminates with a SIGSEGV signal}. The program fails.
-	\end{enumerate}
+		\end{itemize}
+		}}
+		\end{overpic}
+
+		\end{column}
+	\begin{column}{0.6\textwidth}
+		Zeller's prototypical \textit{HOWCOME} algorithm uses DD on program states as described above. When applied to \texttt{fail.c} the Cause-Effect-Chain to the left is returned. \\[2em]
+
+		%The Cause-Effect-Chain makes the reason for \texttt{fail.c} crashing the program obvious: \\
+
+	\end{column}
+	\end{columns}
 
 \end{frame}
+
+\begin{frame}{\textit{HOWCOME} on \texttt{fail.c}}
+	\begin{columns}
+		\begin{column}{.5\textwidth}
+			\begin{overpic}[width=\textwidth]{../figures/check-end.png}
+		\put (0,105) { \parbox{0.9\textwidth}{
+
+			\small
+		\begin{itemize}
+			\item Execution reaches \textbf{if\_then\_else\_cond (95th hit)}.
+\\ Since $*first\_loop\_store\_insn\cecarrow fld[1].rtx \cecarrow fld[1].rtx \cecarrow fld[3].rtx \cecarrow fld[1].rtx$ was $\langle new rtx\_def\rangle$, variable \only<3>{\textbf{$link \cecarrow fld[0].rtx \cecarrow fld[0].rtx$} is now \textbf{$link$}} \only<2>{\red{\textbf{$link \cecarrow fld[0].rtx \cecarrow fld[0].rtx$} is now \textbf{$link$}}}.
+			\item Execution ends.
+Since variable \only<1-2>{$link \cecarrow fld[0].rtx \cecarrow fld[0].rtx$ was $link$, the program now \textbf{terminates with a SIGSEGV signal}}\only<3>{\red{$link \cecarrow fld[0].rtx \cecarrow fld[0].rtx$ was $link$, the program now \textbf{terminates with a SIGSEGV signal}}}.
+The program fails.
+		\end{itemize}
+		}}
+		\end{overpic}
+		\end{column}
+
+		\begin{column}{.5\textwidth}
+			It is now easy to see for the developer why \texttt{fail.c} crashes GCC: \\[1em]
+
+			\onslide<2->{\begin{enumerate}
+				\only<2>{\item A \red{cycle in the RTL tree} of the GCC compiler is created at the 95th hit of \textbf{if\_then\_else\_cond}.}
+				\onslide<3>{\item A cycle in the RTL tree of the GCC compiler is created at the 95th hit of \textbf{if\_then\_else\_cond}.} 
+				\onslide<3>{\item The \red{cycle produces the error}.}
+			\end{enumerate}}
+		\end{column}
+	\end{columns}
+\end{frame}

### 2018-12-18 Vortrag/slides/sections/cause-effect-chains.tex
Extended notes. Minor changes to slides.

@@
 	\begin{column}{0.6\textwidth}
-		Zeller's prototypical \textit{HOWCOME} algorithm uses DD on program states as described above. When applied to \texttt{fail.c} the Cause-Effect-Chain to the left is returned. \\[2em]
 
-		%The Cause-Effect-Chain makes the reason for \texttt{fail.c} crashing the program obvious: \\
-
+		\begin{itemize}
+			\item Zeller's prototypical \textit{HOWCOME} algorithm uses DD on program states as described above.
+			\item When applied to \texttt{fail.c} the Cause-Effect-Chain to the left is returned.
+		\end{itemize}		
+		\phantom{ph}
+		\bigskip
+		\phantom{ph}
 	\end{column}

### 2018-12-18 Vortrag/slides/sections/difficulties.tex
Extended notes. Minor changes to slides.

@@
 \begin{frame}{Difficulty \hashtag 1: Interference}
-	\begin{itemize}
-		\item $ddsimple$ works for single failure inducing-changes.
-			\begin{itemize}
-				\item In each recursion step it applies only the set of changes known to contain a failure inducing change.
-			\end{itemize}
-		\item But what if two changes exist that individually pass the tests but their combination induces failure?
-	\end{itemize}
+	%\begin{itemize}
+	%	\item $ddsimple$ works for single failure inducing-changes.
+	%		\begin{itemize}
+	%			\item In each recursion step it applies only the set of changes known to contain a failure inducing change.
+	%		\end{itemize}
+	%	\item But what if two changes exist that individually pass the tests but their combination induces failure?
+	%\end{itemize}
 
@@
 
-\end{frame}
+	\bigskip
+	\pause
 
-\begin{frame}{Difficulty \hashtag 1: Interference}
 	\begin{exampleblock}{Idea: Leave one set of changes applied}
@@
 
-	\bigskip
+\end{frame}
+
+\begin{frame}{The dd algorithm}
 
 	%If we add this case to the $simpledd$ algorithm it leaves us with Andreas Zellers $dd$ algorithm: $dd(c: 2^{\C}) = dd_2(c,\emptyset) \text{, where}$\\[1em]
+
+	$dd = dd_2(\C, \emptyset)$ \\[1.5em]
 
@@
 		\State let $c_1,c_2 \subseteq c \text{ with } c_1 \cup c_2 = c, c_1 \cap c_2 = \emptyset, |c_1| \approx |c_2|$
+		\State $d_1 = dd_2(c_1, c_2 \cup r)$ %\Comment(``Search in $c_1$ while leaving $c_2$ applied'')
+		\State $d_2 = dd_2(c_2, c_1 \cup r)$ \medskip %\Comment(``Search in $c_2$ while leaving $c_1$ applied'') 
 		\State \Return $ \begin{cases}
 				c & \text{if } |c|=1, \\
-				dd_2(c_1,r) & \text{if } test(c_1 \cup r) = \xmark,\\
-				dd_2(c_2,r) & \text{if } test(c_2 \cup r) = \xmark, \\
-				dd_2(c_1, c_2 \cup r) \cup dd_2(c_2, c_1 \cup r) & \text{otherwise} 
+				dd_2(c_1,r) & \text{if } test(c_1 \cup r) = \xmark, \text{\Comment(``Found in $c_1$'')}\\
+				dd_2(c_2,r) & \text{if } test(c_2 \cup r) = \xmark, \text{\Comment(``Found in $c_2$'')}\\
+				d_1 \cup d_2 & \text{otherwise } \text{\Comment(``Interference'')}
 			\end{cases}$
@@
 \begin{frame}{Difficulty \hashtag 2: Inconsistency}
-	\begin{itemize}
-		\item $dd$ combines changes arbitrarily (in the case of interference)
-		\item This can lead to inconsistent configurations, i.e. no test outcome can be determined for these configurations 
-	\end{itemize}
+	%\begin{itemize}
+	%	\item $dd$ combines changes arbitrarily (in the case of interference)
+	%	\item This can lead to inconsistent configurations, i.e. no test outcome can be determined for these configurations 
+	%\end{itemize}
 
@@
 	\bigskip
+
+	\pause
 	\begin{exampleblock}{Idea: More granular subsets}

### 2018-12-20 Vortrag/slides/sections/ddplus.tex
More notes and nicer use case sections.

--- 2018-11-19 Vortrag/slides/sections/ddplus.tex
@@
 \subsection{The dd+ algorithm}
-\begin{frame}{The dd+ algorithm}
-	\begin{center}
-		Solves it all!
-	\end{center}
+
+\newcommand{\ddp}{$dd^{+}$}
+
+\begin{frame}{The \ddp\ algorithm}
+	Zeller's \textbf{extended} $dd$ algorithm \ddp\ deals with the following cases:
+	\begin{enumerate}
+		\item ``found (in $c_i$)'' \only<2->{$\Rightarrow$ continue search in $c_i$}
+		\item ``interference'' \only<3->{$\Rightarrow$ search in $c_i$ while leaving $\bar{c_i}$ applied and vice versa}
+		\item ``preference'' \only<4->{$\Rightarrow$ search in $c_i$ while leaving $\bar{c_i}$ applied}
+		\item ``try again'' \only<5->{$\Rightarrow$ repeat search on same change set with twice as many subsets}
+		%\item ``nothing left		
+	\end{enumerate}
 \end{frame}
+
+\begin{frame}{The \ddp\ algorithm}
+	Properties of \ddp:
+	\begin{itemize}
+		\item \ddp finds a minimal failure-inducing subset of changes as long as the changes are safe, i.e. combined they result in a consistent configuration.
+		\item \ddp has at most linear time complexity like $dd$.
+	\end{itemize}
+\end{frame}

### 2018-12-20 Vortrag/figures/fail.c

@@ -1,0 +1,12 @@
+	
+
+	double mult(double z[], int n)
+	{
+		int i,j;
+		i = 0;
+		for (j = 0; j < n; j++) {
+			i = i + j + 1;
+			z[i] = z[i] * (z[0] + 1.0);
+		}
+		return z[n];
+	}

### 2018-12-20 Paper/article.tex
Changed paper to the LIPIcs standard.

\documentclass[a4paper,UKenglish]{lipics-v2018}

\usepackage{microtype}

\bibliographystyle{plainurl}

\title{A summary on Delta Debugging and its uses}

\author{Moritz C. Laupichler}{Fakultät für Informatik, Karlsruhe Institute of Technology, Germany}{moritz.laupichler@student.kit.edu}{}{}

\authorrunning{M.\, C. Laupichler}

\Copyright{Moritz C. Laupichler}

\subjclass{\ccsdesc[500]{Software and its engineering~Software testing and debugging}}

\keywords{Delta Debugging}

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{Moritz Laupichler}
\EventNoEds{1}
\EventLongTitle{Proseminar Werkzeuge und Methoden der Software-Analyse, WS18/19}
\EventShortTitle{PS WMSA WS18/19}
\EventAcronym{WMSA}
\EventYear{2018}
\EventDate{October 19, 2018 -- February 1, 2019}
\EventLocation{Karlsruhe, Germany}
\EventLogo{}
\SeriesVolume{1}
\ArticleNo{M6}
%\nolinenumbers %uncomment to disable line numbering
%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\nolinenumbers
\hideLIPIcs

\begin{document}

\maketitle

\begin{abstract}
	This paper deals with the Delta Debugging method.
\end{abstract}

\section{What is Delta Debugging?}
Lelokek
\subsection{Yesterday and today}
\subsection{The simple idea}

\section{Definitions and basework}
\subsection{Configurations}
\subsection{Problems: Interference and Inconsistency}

\section{The dd+ algorithm}
\subsection{How to solve interference}
\subsection{How to solve inconsistency}
\subsection{The dd+ algorithm and its properties}

\section{Using DD to minimize random unit test cases}
\subsection{The ddmin algorithm}
\subsection{A modern approach}

\end{document}

### 2018-12-29 Paper/article.tex
Began actual work on the Paper.

@@
 \ArticleNo{M6}
-%\nolinenumbers %uncomment to disable line numbering
-%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \nolinenumbers
 \hideLIPIcs
+
+%%%%%%%% Own Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Special symbol font
+\usepackage{pifont}
+\Pifont{pzd}
+
+% Pseudocode algorithms
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+\algrenewcommand{\algorithmiccomment}[1]{\hfill\(//\) #1}
+
+%%%%%%%%%%%% Custom Commands %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% DD algorithms abbreviation
+\newcommand{\sdd}[0]{\textit{simpledd}}
+\newcommand{\dd}[0]{\textit{dd}}
+\newcommand{\ddp}{\textit{dd\textsuperscript{+}}}
+
+% Colored text shortcuts
+\newcommand{\green}[1]{\textcolor{green}{#1}}
+\newcommand{\red}[1]{\textcolor{red}{#1}}
+\newcommand{\gray}[1]{\textcolor{lightgray}{#1}}
+
+% Yesterday and Today
+\newcommand{\yd}[0]{\green{Yesterday} }
+\newcommand{\td}[0]{\red{Today} }
+
+% The global change set C
+\newcommand{\C}[0]{\ensuremath{\mathcal{C}}}
+
+% Possible results of the test function
+\newcommand{\cmark}{\text{\ding{51}}}
+\newcommand{\xmark}{\text{\ding{55}}}
+\newcommand{\qmark}{\textbf{?}}
+
+% Definition subject
+\newcommand{\defsub}[1]{\textbf{(#1)} }
+
+% Math mode shortcuts
+\newcommand{\set}[1]{\left \{ #1 \right \}}
 
@@
 \section{What is Delta Debugging?}
-Lelokek
+
 \subsection{Yesterday and today}
+\label{ydandtd}
+
+Imagine a software development process where the latest changes to the code cause known tests to fail even though an older version of the code passed these exact tests.
+As the basis of the Delta Debugging principle we need to define a simple abstraction of this point in the development process: We know a state \yd that passes the tests and a state \td that fails the tests.
+Furthermore we know all changes that when applied to \yd yield \td.
+\\
+The goal of Delta Debugging is to automate extraction of the exact changes to \yd that cause the failure in \td.
+
 \subsection{The simple idea}
+\label{ddidea}
+
+The basic idea of Delta Debugging is similar to a binary search.
+Recursively the number of candidates for changes that cause \td to fail is lowered.
+By disregarding the changes that do not cause failure eventually we end up with the smallest set of changes that when applied to \yd causes the tests to fail.
 
 \section{Definitions and basework}
+
+The ideas described above need to be abstracted and formalized in order to approach Delta Debugging algorithms.
+
 \subsection{Configurations}
-\subsection{Problems: Interference and Inconsistency}
 
-\section{The dd+ algorithm}
-\subsection{How to solve interference}
-\subsection{How to solve inconsistency}
-\subsection{The dd+ algorithm and its properties}
+\definition \defsub{Configuration} A change set $c \subseteq \C$ is called \textbf{configuration} where $\C$ is the set of all changes between \yd and \td.
 
-\section{Using DD to minimize random unit test cases}
-\subsection{The ddmin algorithm}
-\subsection{A modern approach}
+\definition \defsub{Baseline} The empty configuration $\emptyset$ is called a \textbf{baseline}.
+
+\definition \defsub{Test} The function $test: 2^{\C} \to \set{\cmark, \xmark, \qmark}$ determines the test outcome for a configuration. It returns one of three possible results: 
+\begin{itemize}
+	\item The tests pass: $\cmark$
+	\item The tests fail: $\xmark$
+	\item No result can be determined: $\qmark$
+\end{itemize}
+
+Let $c \subseteq \C$ be a configuration.
+$test(c)$ models a suite of regression tests being run on the result of applying all changes in $c$ to \yd.
+So by definition $test(\emptyset) = \cmark$ ("Yesterday passes") and $test(\C) = \xmark$ ("Today fails") hold.
+
+\definition \defsub{Failure-inducing configuration} A configuration $c$ is called \textbf{failure-inducing} iff 
+\[ \forall c' (c \subseteq c' \subseteq \C \rightarrow test(c') \ne \cmark) \] 
+holds.
+
+\definition \defsub{Minimal failure-inducing configuration} A failure-inducing configuration $c$ is \textbf{minimal} iff 
+\[ \forall c' \subset c: test(c') \ne \xmark \]
+holds.\\
+
+These terms abstract the software development situation described in \ref{ydandtd}. The goal of Delta Debugging can now be stated as finding a minimal failure-inducing configuration in $\C$.
+
+\section{Incrementally approaching a plausible DD algorithm}
+
+There is two major difficulties when approaching a Delta Debugging algorithm: interference and inconistency.
+A na\"ive Delta Debugging algorithm called \sdd will be refined into the \dd algorithm and finally the \ddp algorithm in this chapter in order to describe and solve both difficulties.
+
+\subsection{A na\"ive algorithm: \sdd}
+
+The na\"ive algorithm \sdd\ recursively splits \C\ into smaller and smaller parts until one failure-inducing change is found.
+In each recursion step the current set of changes, $c$, is partitioned in two halves, $c_1$ and $c_2$, and both halves are tested.
+The half that fails the tests is regarded in the next recursion step.
+The algorithm knows two cases, "Found in $c_1$" and "Found in $c_2$".
+\\
+
+\begin{algorithmic}[1]
+		\Function{simpledd}{$c: 2^{\C}$}
+			\If{$|c| = 1$} \Return $c$ \EndIf
+			\State Partition $c$ into two halves $c_1$, $c_2$ so that $c_1 \cap c_2 = \emptyset$
+			\If{($test(c_1) = \xmark$)} \Return $simpledd(c_1)$ \Comment(``Found in $c_1$'') \Else{} \Return $simpledd(c_2)$ \Comment(``Found in $c_2$'')
+			\EndIf
+		\EndFunction
+\end{algorithmic}
+
+It becomes evident that this algorithm is not effective because it is only capable of finding single failure-inducing changes.
+An idea for a solution would be to remove the change that is found and re-run the algorithm.
+But what if mutiple changes pass the test individually but their combination induces failure?
+\\
+
+\definition \defsub{Interference} Two configurations $c_1$ and $c_2$ \textbf{interfere} iff $test(c_1) = \cmark$, $test(c_2) = \cmark$ but $test(c_1 \cup c_2) = \xmark$.
+
+\subsection{Dealing with interference: \dd}
+
+\subsection{Dealing with inconsistency: \ddp}
+
+\section{The \ddp algorithm}
+\subsection{The \ddp algorithm and its properties}
+
+\section{Exemplary use cases}
+\subsection{Program input analysis}
+\subsection{Program runtime analysis}
+\subsection{Minimizing random unit test cases}
+\subsubsection{The ddmin algorithm}
+
+\section{Reflection and conclusion}
+\subsection{Abstractness: strength and weakness}
+\subsection{Conclusion}
 

### 2018-12-29 Paper/article.tex
More additions to paper.

@@
 
-There is two major difficulties when approaching a Delta Debugging algorithm: interference and inconistency.
-A na\"ive Delta Debugging algorithm called \sdd will be refined into the \dd algorithm and finally the \ddp algorithm in this chapter in order to describe and solve both difficulties.
+There is two major difficulties when approaching a Delta Debugging algorithm: interference and inconsistency.
+A na\"ive Delta Debugging algorithm called \sdd\ will be refined into the \dd\ algorithm and finally the \ddp\ algorithm in this chapter in order to describe and solve both difficulties.
 
@@
 			\State Partition $c$ into two halves $c_1$, $c_2$ so that $c_1 \cap c_2 = \emptyset$
-			\If{($test(c_1) = \xmark$)} \Return $simpledd(c_1)$ \Comment(``Found in $c_1$'') \Else{} \Return $simpledd(c_2)$ \Comment(``Found in $c_2$'')
+			\If{($test(c_1) = \xmark$)} \Return $simpledd(c_1)$ \Comment("Found in $c_1$") \Else{} \Return $simpledd(c_2)$ \Comment("Found in $c_2$")
 			\EndIf
@@
 An idea for a solution would be to remove the change that is found and re-run the algorithm.
-But what if mutiple changes pass the test individually but their combination induces failure?
+But what if mutiple changes pass the test individually but only their combination induces failure?
 \\
@@
 
+\sdd\ is unable to handle the case of both $c_1$ and $c_2$ passing.
+A new algorithm \dd\ is created which behaves like \sdd\ but incorporates the case "Interference".
+In this case the algorithm searches for the failure-inducing change in $c_1$ while keeping $c_2$ applied and vice versa.
+By combining the results of both searches the interfering changes can be extracted.
+\\[.5em]
+
+$dd = dd_2(\C, \emptyset)$ \\
+
+\begin{algorithmic}[1]
+	\Function{$dd_2$}{$c,r:2^\C$} $: 2^{\C}$
+	\State let $c_1,c_2 \subseteq c \text{ with } c_1 \cup c_2 = c, c_1 \cap c_2 = \emptyset, |c_1| \approx |c_2|$
+	\State $d_1 = dd_2(c_1, c_2 \cup r)$ \Comment("Search in $c_1$ while leaving $c_2$ applied")
+	\State $d_2 = dd_2(c_2, c_1 \cup r)$ \Comment("Search in $c_2$ while leaving $c_1$ applied")\medskip  
+	\State \Return $ \begin{cases}
+			c & \text{if } |c|=1, \\
+			dd_2(c_1,r) & \text{if } test(c_1 \cup r) = \xmark, \text{\Comment("Found in $c_1$")}\\
+			dd_2(c_2,r) & \text{if } test(c_2 \cup r) = \xmark, \text{\Comment("Found in $c_2$")}\\
+			d_1 \cup d_2 & \text{otherwise } \text{\Comment("Interference")}
+		\end{cases}$
+	\EndFunction
+\end{algorithmic}
+
+In $dd_2$ the parameter $c$ describes the configuration in which the algorithm is currently searching and the parameter $r$ contains all changes that remain applied during the search in $c$. \\
+
+In \dd\ changes are combined arbitrarily and then tested. Depending on the changes it is possible for various reasons like syntactical errors %TODO more reasons
+that tests on the resulting configurations can not be resolved.
+
+\definition \defsub{Inconsistent configuration} A configuration $c \subseteq \C$ is called \textbf{inconsistent} iff $test(c) = \qmark$.\\
+
+\dd\ does not deal with inconsistent configurations.
+
 \subsection{Dealing with inconsistency: \ddp}
 
-\section{The \ddp algorithm}
-\subsection{The \ddp algorithm and its properties}
+\section{The \ddp\ algorithm}
+\subsection{The \ddp\ algorithm and its properties}
 

### 2018-12-30 Paper/article.tex
More work on the paper (up to dd+)

@@
 \newcommand{\set}[1]{\left \{ #1 \right \}}
+\newcommand{\nN}[0]{\mathbb{N}}
 
@@
 \\
+\begin{figure}[h!]
+\fbox{\begin{minipage}{\textwidth}
 
@@
 \end{algorithmic}
-
-It becomes evident that this algorithm is not effective because it is only capable of finding single failure-inducing changes.
-An idea for a solution would be to remove the change that is found and re-run the algorithm.
+\end{minipage}}
+\caption{The na\"ive DD algorithm \sdd}
+\end{figure}
+\\
+It becomes evident that this algorithm is not effective because it is only capable of finding single changes that cause failure on their own.
 But what if mutiple changes pass the test individually but only their combination induces failure?
@@
 
-\definition \defsub{Interference} Two configurations $c_1$ and $c_2$ \textbf{interfere} iff $test(c_1) = \cmark$, $test(c_2) = \cmark$ but $test(c_1 \cup c_2) = \xmark$.
+\definition \defsub{Interference} Two configurations $c_1$ and $c_2$ \textbf{interfere} iff $test(c_1) = \cmark$, $test(c_2) = \cmark$ but $test(c_1 \cup c_2) = \xmark$. \\
 
@@
 By combining the results of both searches the interfering changes can be extracted.
-\\[.5em]
+\\
+\begin{figure}[h!]
+\fbox{\begin{minipage}{\textwidth}
 
-$dd = dd_2(\C, \emptyset)$ \\
+$dd = dd_2(\C, \emptyset)$ where \\
 
@@
 \end{algorithmic}
+\end{minipage}}
+\caption{The \dd\ algorithm capable of dealing with interference.}
+\end{figure}
+\\
 
@@
 
-In \dd\ changes are combined arbitrarily and then tested. Depending on the changes it is possible for various reasons like syntactical errors %TODO more reasons
-that tests on the resulting configurations can not be resolved.
+In \dd\ changes are combined arbitrarily and then tested. It is possible that no test result on the resulting configurations can be determined.
 
-\definition \defsub{Inconsistent configuration} A configuration $c \subseteq \C$ is called \textbf{inconsistent} iff $test(c) = \qmark$.\\
+\definition \defsub{Inconsistent configuration} A configuration $c \subseteq \C$ is called \textbf{inconsistent} iff $test(c) = \qmark$. \\
 
-\dd\ does not deal with inconsistent configurations.
+In the real world of code and regression tests this scenario is most likely caused by a combination of changes resulting in syntactical or semantical errors.
+An inconsistent configuration may also occur when a change is dependent on another change not included in the configuration or when the program execution fails during testing.
+
+\dd\ does not deal with inconsistent configurations. It behaves wrong if the configurations that are tested in the "Found" cases are inconsistent.
 
 \subsection{Dealing with inconsistency: \ddp}
+
+The key to solving inconsistency is partitioning $c$ into more than two subsets making the sets of changes smaller.
+We know that the baseline $\emptyset$ and applying all changes $\C$ are consistent.
+By applying less changes to \yd or removing less changes from \td (s.
+case "Preference" below).
+the chances of the result being consistent is increased.
+If the algorithm runs into only inconsistent configurations the number of subsets is increased and the chances of at least one configuration being consistent after is increased.
+This solution requires the final Delta Debugging algorithm to handle any number $n$ of subsets whereas \dd\ only handles two change sets, $c_1$ and $c_2$ at a time.
+
+\definition \defsub{Complement of configuration} Let $c \subseteq \C$ be a configuration. The configuration $\bar{c} = \C \setminus c$ is called the \textbf{complement} of $c$.\\
+
+The final Delta Debugging algorithm \ddp\ on the subsets $c_1, \dots, c_n$ knows four cases:\\
+\begin{description}
+	\item["Found in $c_i$"] If testing any $c_i$ fails it contains a failure-inducing subset. Continue search in $c_i$.
+	\item["Interference"] If any $c_i$ and $\bar{c_i}$ both pass they form an interference. Continue as in \dd.
+	\item["Preference"] If any $c_i$ is unresolved but $\bar{c_i}$ passes then $c_i$ contains a failure-inducing subset.
+$c_i$ is then preferred.
+Continue search in $c_i$ while leaving $\bar{c_i}$ applied to promote consistency.
+	\item["Try Again"] If none of the above cases apply then continue the search with $2n$ subsets. The chance of consistent configurations is increased.
+\end{description}
+
+\begin{figure}[h!]
+\fbox{\begin{minipage}{\textwidth}
+
+$dd^+(c) = dd_3(c, \emptyset,2)$ where \\
+
+\begin{algorithmic}[1]
+	\Function{$dd_3$}{$c,r:2^\C, n: \nN$} $: 2^{\C}$
+	\State let $c_1,\dots,c_n \subseteq c \text{ with } \bigcup c_i = c, c_i \text{ pairwise disjoint}, \forall i: |c_i| \approx |c|/n$
+	\State let $\bar{c_i} = c \setminus (c_i \cup r), t_i = test(c_i \cup r), \bar{t_i} = test(\bar{c_i} \cup r)$
+	\State let $c' = c \cap \bigcap \set{\bar{c_i} \mid \bar{t_i} = \xmark}, r' = r \cup \bigcup \set{c_i \mid t_i = \cmark}, n' = min (\mid c' \mid, 2n)$
+	\State let $d_i = dd_3(c_i, \bar{c_i} \cup r, 2), \bar{d_i} = dd_3(\bar{c_i},c_i \cup r, 2)$ 
+	\State \Return $ \begin{cases}
+			c & \text{if } |c|=1, \text{\Comment("Found")}\\
+			dd_3(c_i, r, 2) & \text{if } t_i = \xmark \text{ for some i \Comment("Found in $c_i$")}\\
+			d_i \cup \bar{d_i} & \text{if } t_i = \cmark \wedge \bar{t_i} = \cmark \text{ for some i \Comment("Interference")}\\
+			d_i & \text{if } t_i = \qmark \wedge \bar{t_i} = \cmark \text{ for some i \Comment("Preference")}\\
+			dd_3(c',r',n') & \text{if } n < \mid c \mid \text{\Comment("Try Again")}\\
+			c' & \text{otherwise \Comment("nothing left")}
+		\end{cases}$
+	\EndFunction
+\end{algorithmic}
+\end{minipage}}
+\caption{The \ddp\ algorithm capable of dealing with unresolved test cases.}
+\end{figure}
 

### 2019-01-02 Paper/article.tex
Began use cases in the paper.

@@
 \algrenewcommand{\algorithmiccomment}[1]{\hfill\(//\) #1}
+
+% Colors
+\usepackage{xcolor}
 
@@
 
-\section{The \ddp\ algorithm}
-\subsection{The \ddp\ algorithm and its properties}
+\section{Exemplary use cases}
 
-\section{Exemplary use cases}
+In this chapter three exemplary use cases for Delta Debugging will be shown.
+
 \subsection{Program input analysis}
+
+If a certain input is known to cause a program to fail execution Delta Debugging can be used to analyze which part of the program input is responsible for the failure.
+The \ddp\ algorithm as defined above can be used without any modifications.
+It is necessary, though, to define certain abstract concepts in the context of input analysis.\\
+\begin{itemize}
+  	\item We assume a known minimal input for the program at hand that does not cause failure.
+In many cases this minimal input would be some sort of "empty" input, be it no data, zero values, etc.
+This minimal input is regarded as the \yd state.
+  	\item The known failing input is regarded as the \td state.
+  	\item A change is defined as the smallest possible difference between any two possible inputs, usually a letter, number or other token of data.
+$\C$ can then be any set of changes that convert the minimal input \yd to the failing input \td.
+  	\item The $test$ function returns $\xmark$ if exactly the known error occurs on program execution. It returns $\cmark$ if no error occurs and $\qmark$ in all other cases. 
+\end{itemize}
+
+Running \ddp\ with these concrete definitions finds the parts of the input that are responsible for the failure.
+If the input itself is faulty the error will be revealed directly.
+If the error lies within the program the algorithm delivers a starting point for manually debugging the error.
+
+Consider as an example the C program \texttt{fail.c} which caused the GNU C Compiler (GCC) in version XY %TODO VERSION!
+to crash.
+The program looks like correct C code that should be able to be compiled.
+We'd like to use Delta Debugging to find out which part of \texttt{fail.c} is responsible for the crash.
+For this an empty C program is regarded as the \yd state and \texttt{fail.c} as the \td state.
+A change is defined as a C language token.
+$\C$ is comprised of all tokens in \texttt{fail.c}.
+As depicted in \ref{fig:ddponfail} \ddp\ only takes 21 tests to find out that the C tokens "$+ 1.0$" are a minimal failure-inducing set of changes.
+\\
+\begin{figure}[h!]
+	\fbox{
+		\begin{minipage}{\textwidth}
+			\begin{centering}
+			\begin{tabular}{clc}
+				\# & GCC input & test \\
+				\hline
+				1 & double mult($\dots$) $\{ int \; i, j; \; i = 0; \; for (\dots) \{ \dots\} \dots\}$ & $\xmark$ \\
+				2 &\gray{ double mult($\dots$) $\{ int \; i, j; \; i = 0; \; for (\dots) \{ \dots\} \dots\}$} & $\cmark$ \\
+				3 & double mult($\dots$) $\{\gray{ int \; i, j; \; i = 0; \; for (\dots) \{ \dots\} \dots}\}$ & $\cmark$ \\
+				4 & double mult($\dots$) $\{ int \; i, j; \; i = 0; \; \gray{for (\dots) \{ \dots\} \dots}\}$ & $\cmark$ \\
+				5 & double mult($\dots$) $\{ int \; i, j; \; i = 0; \; for (\dots) \{ \dots\}\gray{ \dots}\}$ & $\xmark$ \\
+				6 & double mult($\dots$) $\{ int \; i, j; \; i = 0; \; for (\gray{\dots}) \{ \dots\} \gray{\dots}\}$ & $\cmark$ \\
+
+				\vdots & \multicolumn{1}{c}{\vdots} & \vdots \\
+				18 & \multicolumn{1}{c}{\dots \quad $z[i] = z[i] * (z[0] + 1.0);$ \quad \dots } & \xmark \\
+				19 & \multicolumn{1}{c}{\dots \quad $z[i] = z[i] * (z[0] \gray{+ 1.0});$ \quad \dots } & \cmark \\
+				20 & \multicolumn{1}{c}{\dots \quad $z[i] = z[i] * (z[0] \gray{+} 1.0);$ \quad \dots } & \qmark \\
+				21 & \multicolumn{1}{c}{\dots \quad $z[i] = z[i] * (z[0] + \gray{1.0});$ \quad \dots } & \qmark \\
+			\end{tabular}
+			\end{centering}
+		\end{minipage}
+	}
+	\label{fig:ddponfail}
+	\caption{How \ddp\ finds the part of \texttt{fail.c} that is responsible for GCC crashing.}
+\end{figure}
+
 \subsection{Program runtime analysis}

### 2019-01-02 Paper/article.tex
Minor fixes in the algorithms chapter.

--- 2018-12-30 Paper/article.tex
@@
 
-\definition \defsub{Complement of configuration} Let $c \subseteq \C$ be a configuration. The configuration $\bar{c} = \C \setminus c$ is called the \textbf{complement} of $c$.\\
+\definition \defsub{Complement of configuration} Let $c \subseteq \C$ be a configuration.
+The configuration $\bar{c}$ is called the \textbf{complement} of $c$.
+It contains all changes that are contained in the current search but not in $c$.\\
 
@@
 			dd_3(c',r',n') & \text{if } n < \mid c \mid \text{\Comment("Try Again")}\\
-			c' & \text{otherwise \Comment("nothing left")}
+			c' & \text{otherwise \Comment("Nothing Left")}
 		\end{cases}$

### 2019-01-11 Paper/article.tex
Started on use case 2.

@@
 \usepackage{xcolor}
+\definecolor{td-green}{rgb}{0,.59,.51}
+
+% Better references
+\usepackage{hyperref}
 
@@
 % Colored text shortcuts
-\newcommand{\green}[1]{\textcolor{green}{#1}}
+\newcommand{\green}[1]{\textcolor{td-green}{#1}}
 \newcommand{\red}[1]{\textcolor{red}{#1}}
@@
 \newcommand{\nN}[0]{\mathbb{N}}
+
+% fail.c reference shortcut
+\newcommand{\reffail}[0]{\href{fig:fail.c}{\ensuremath{fail.c^{Fig. \ref{fig:fail.c}}}}}
+
+% r_pass and r_fail shortcuts
+\newcommand{\rpass}[0]{\ensuremath{r_{\cmark}}}
+\newcommand{\rfail}[0]{\ensuremath{r_{\xmark}}}
 
@@
 
-Consider as an example the C program \texttt{fail.c} which caused the GNU C Compiler (GCC) in version XY %TODO VERSION!
-to crash.
-The program looks like correct C code that should be able to be compiled.
-We'd like to use Delta Debugging to find out which part of \texttt{fail.c} is responsible for the crash.
-For this an empty C program is regarded as the \yd state and \texttt{fail.c} as the \td state.
+Consider as an example the C program \reffail which caused the GNU C Compiler (GCC) in version 2.95.2 (on Intel-Linux with optimization enabled) to crash.
+The program looks like correct C code that should be compiled properly.
+We'd like to use Delta Debugging to find out which part of \reffail is responsible for the crash.
+An empty C program is regarded as the \yd state and \reffail as the \td state.
 A change is defined as a C language token.
-$\C$ is comprised of all tokens in \texttt{fail.c}.
+$\C$ is comprised of all tokens in \reffail.
 As depicted in \ref{fig:ddponfail} \ddp\ only takes 21 tests to find out that the C tokens "$+ 1.0$" are a minimal failure-inducing set of changes.
+\\
+\begin{figure}[h!]
+	\fbox{
+		\begin{minipage}{\textwidth}
+		\begin{centering}
+			\begin{algorithmic}[1]
+					\State double \textbf{mult}(double z[], int n)
+					\State $\{$
+					\State \quad int i,j;
+					\State \quad i = 0;
+					\State \quad for(j = 0; j < n; j++) $\{$
+					\State \quad \quad i = i + j + 1;
+					\State \quad \quad z[i] = z[i] * (z[0] + 1.0);
+					\State \quad $\}$
+					\State \quad return z[n];
+					\State $\}$
+			\end{algorithmic}
+		\end{centering}
+		\end{minipage}
+	}
+	\caption{The $fail.c$ program that crashes GCC}
+	\label{fig:fail.c}
+\end{figure}
 \\
@@
 	}
+	\caption{How \ddp\ finds the part of \texttt{fail.c} that is responsible for GCC crashing.}
 	\label{fig:ddponfail}
-	\caption{How \ddp\ finds the part of \texttt{fail.c} that is responsible for GCC crashing.}
 \end{figure}
@@
 \subsection{Program runtime analysis}
+
+It can be of great help to the software developer to be able to automatically determine the specific parts of the program input that cause the failure.
+The information gathered from the above method can be used by the developer to manually trace the responsible input data to the crash.
+Considering how a miniscule input change can cause entirely different execution paths this task is tedious if not impossible.
+Delta Debugging can help trace the error from the root to the failure during program execution as well, though.
+Given a test framework that is able to access program states during execution, DD is used on the program states of two program runs \rpass\ and \rfail\ at significant locations to deduce a \textit{Cause-Effect-Chain (CEC)}.
+
+\definition \defsub{Program states} A program state
+
+The entire program execution is determined by a set of circumstances.
+We focus on changeable circumstances, more specifically the circumstances that influence the test outcome.
+A program run, i.e.
+one specific sequence of program states, is determined only by the set of circumstances that induce the specific run.
+
+\definition \defsub{Circumstances and program runs} Let $R$ be the set of all possible circumstance configurations.
+A run is defined as a configuration of circumstances, so every $r \in R$ is called a run.
+
 \subsection{Minimizing random unit test cases}

### 2019-01-13 Paper/article.tex
Lots of work on dddiff and test case minimization.

@@
 
-% Better references
-\usepackage{hyperref}
-
 %%%%%%%%%%%% Custom Commands %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \newcommand{\ddp}{\textit{dd\textsuperscript{+}}}
+\newcommand{\dddiff}{\textit{dd\textsuperscript{diff}}}
 
@@
 % Yesterday and Today
-\newcommand{\yd}[0]{\green{Yesterday} }
-\newcommand{\td}[0]{\red{Today} }
+\newcommand{\yd}[0]{\green{Yesterday}}
+\newcommand{\td}[0]{\red{Today}}
 
@@
 \newcommand{\qmark}{\textbf{?}}
+
+% Failing and passing configuration
+\newcommand{\cpass}{\ensuremath{c_{\cmark}}}
+\newcommand{\cfail}{\ensuremath{c_{\xmark}}}
 
@@
 \newcommand{\rfail}[0]{\ensuremath{r_{\xmark}}}
+\newcommand{\spass}[0]{\ensuremath{s_{\cmark}}}
+\newcommand{\sfail}[0]{\ensuremath{s_{\xmark}}}
+
+% Cause-Effect-Chain arrow
+\newcommand{\cecarrow}{\to}
 
@@
 Imagine a software development process where the latest changes to the code cause known tests to fail even though an older version of the code passed these exact tests.
-As the basis of the Delta Debugging principle we need to define a simple abstraction of this point in the development process: We know a state \yd that passes the tests and a state \td that fails the tests.
-Furthermore we know all changes that when applied to \yd yield \td.
+As the basis of the Delta Debugging principle we need to define a simple abstraction of this point in the development process: We know a state \yd\ that passes the tests and a state \td\ that fails the tests.
+Furthermore we know all changes that when applied to \yd\ yield \td.
 \\
-The goal of Delta Debugging is to automate extraction of the exact changes to \yd that cause the failure in \td.
+The goal of Delta Debugging is to automate extraction of the exact changes to \yd\ that cause the failure in \td.
 
@@
 The basic idea of Delta Debugging is similar to a binary search.
-Recursively the number of candidates for changes that cause \td to fail is lowered.
-By disregarding the changes that do not cause failure eventually we end up with the smallest set of changes that when applied to \yd causes the tests to fail.
+Recursively the number of candidates for changes that cause \td\ to fail is lowered.
+By disregarding the changes that do not cause failure eventually we end up with the smallest set of changes that when applied to \yd\ causes the tests to fail.
 
@@
 
-\definition \defsub{Configuration} A change set $c \subseteq \C$ is called \textbf{configuration} where $\C$ is the set of all changes between \yd and \td.
+\definition \defsub{Configuration} A change set $c \subseteq \C$ is called \textbf{configuration} where $\C$ is the set of all changes between \yd\ and \td.
 
@@
 There is two major difficulties when approaching a Delta Debugging algorithm: interference and inconsistency.
-A na\"ive Delta Debugging algorithm called \sdd\ will be refined into the \dd\ algorithm and finally the \ddp\ algorithm in this chapter in order to describe and solve both difficulties.
+A na\"ive Delta Debugging algorithm called \sdd\ will be refined into the \dd\ algorithm and the \ddp\ algorithm in this chapter in order to describe and solve both difficulties.
+Finally, isolation of a minimal-failure inducing difference, an alternative approach to DD, is protrayed through the \dddiff\ algorithm.
 
@@
 \caption{The na\"ive DD algorithm \sdd}
+\label{fig:sdd}
 \end{figure}
@@
 \caption{The \dd\ algorithm capable of dealing with interference.}
+\label{fig:dd}
 \end{figure}
@@
 We know that the baseline $\emptyset$ and applying all changes $\C$ are consistent.
-By applying less changes to \yd or removing less changes from \td (s.
+By applying less changes to \yd\ or removing less changes from \td\ (s.
 case "Preference" below).
@@
 \caption{The \ddp\ algorithm capable of dealing with unresolved test cases.}
+\label{fig:ddp}
 \end{figure}
+
+\subsection{Isolating a minimal difference: \dddiff}
+
+All the above Delta Debugging algorithms have focused on simplification of a failing configuration.
+The idea of simplification is to eliminate as many non-failure-inducing changes from $\C$ as possible to find out what the actual changes between \yd\ and \td\ were that cause the failure today.
+
+Another goal that can be achieved by Delta Debugging is the isolation of the minimal failure-inducing difference between \yd\ and \td.
+Isolation means to replicate this difference through two other configurations \cpass\ and \cfail\ that will not not make any sense from a production standpoint but portray the difference in a minimal and thus simple to understand fashion.
+This is exactly what the programmer wants when debugging the code: A simplistic reconstruction of what separates passing code from the known failing code in the given context, i.e.
+the reason why \td\ fails.\\
+
+In more technical terms the goal is to find two configurations \cpass and \cfail, such that $\emptyset \subseteq \cpass \subset \cfail \subseteq \C$, $test(\cpass) = \cmark$ and $test(\cfail) = \xmark$ hold and the difference between the configurations $\Delta = \cfail \setminus \cpass$ is minimal.
+
+\definition \label{def:minimal_difference} \defsub{Minimal failure-inducing difference} Let \cpass\ and \cfail\ be two test cases with $\emptyset \subseteq \cpass \subset \cfail \subseteq \C$.
+Their difference $\Delta = \cfail \setminus \cpass$ is \textbf{minimal} iff \[ \forall \Delta_i \subset \Delta : test(\cpass \cup \Delta_i) \ne \cmark \wedge test(\cfail \setminus \Delta_i) \ne \xmark \]
+
+The number of subsets of $\Delta$ is exponential so we relax the definition of minimality to n-minimality.
+
+\definition \label{def:n-minimal_difference} \defsub{N-minimal failure-inducing difference} Let \cpass\ and \cfail\ be defined as above.
+Their difference $\Delta = \cfail \setminus \cpass$ is \textbf{n-minimal} iff \[ \forall \Delta_i \subset \Delta : |\Delta_i| \le n \Rightarrow test(\cpass \cup \Delta_i) \ne \cmark \wedge test(\cfail \setminus \Delta_i) \ne \xmark \]
+
+We aim at finding two configurations with a \textit{1-minimal failure-inducing difference} - that is a configuration \cfail\ where taking away any change $\delta \in \Delta$ will make the failure disappear.\\
+
+\begin{figure}[t]
+	\fbox{
+		\begin{minipage}{\textwidth}
+		The goal of \dddiff\ is to find $(\cfail, \cpass) = \dddiff(\C)$ such that $\emptyset \subseteq \cpass \subset \cfail \subseteq \C$, $test(\cpass) = \cmark$, $test(\cfail) = \xmark$, and $\Delta = \cfail \setminus \cpass$ is 1-minimal.\\
+
+		The \dddiff\ algorithm is $\dddiff(\C)=dd_4(\emptyset, \C, 2)$ where
+
+				\begin{algorithmic}[1]
+					\Function{$dd_4$}{$\cpass,\cfail:2^{\C}$, $n: \nN$}: $2^{\C} \times 2^{\C}$
+					\State let $\Delta = \cfail \setminus \cpass$
+					\State let $\Delta = \bigcup^{n}_{i=1} \Delta_i$ with all $\Delta_i$ pairwise disjoint and $\forall \Delta_i: |\Delta_i| \approx |\Delta|/n$
+					\State let $t_{\cmark, i} = test(\cpass \cup \Delta_i)$ for $i \in \set{1, \dots, n}$ \Comment(``test for larger passing configuration'')
+					\State let $t_{\xmark, i} = test(\cfail \setminus \Delta_i)$ for $i \in \set{1, \dots, n}$ \Comment(``test for smaller failing configuration'')
+					\State \Return $\begin{cases}
+						dd_4(\cpass, \cpass \cup \Delta_i,2) & \text{if } \; t_{\cmark, i} = \xmark \text{\Comment(``reduce to subset'')} \\
+						dd_4(\cfail \setminus \Delta_i, \cfail, 2) & \text{if } t_{\xmark, i} = \cmark \text{\Comment(``increase to complement'')}\\
+						dd_4(\cpass \cup \Delta_i, \cfail, max(n-1, 2)) & \text{if } t_{\cmark, i} = \cmark \text{\Comment(``increase to subset'')}\\
+						dd_4(\cpass, \cfail \setminus \Delta_i, max(n-1,2)) & \text{if } t_{\xmark, i} = \xmark \text{\Comment(``reduce to complement'')}\\
+						dd_4(\cpass, \cfail, min(2n, |\Delta|)) & \text{if } n < |\Delta| \text{\Comment(``increase granularity'')} \\
+						(\cpass, \cfail) & \text{otherwise \Comment(``done'')} 
+					\end{cases}$
+					(check the first cases for all $i \in \set{1, \dots, n}$)
+					\EndFunction
+				\end{algorithmic}
+		\end{minipage}
+	}
+	\caption{The Delta Debugging algorithm for finding two configurations with minimal difference, \dddiff.}
+	\label{fig:dddiff}
+\end{figure}
+
+The idea of the isolation algorithm \dddiff\ is to increase the size of \cpass\ or decrease the size of \cfail\ until their difference is 1-minimal.
+The algorithm is initialized with $\cpass = \emptyset$ and $\cfail = \C$.
+The difference $\Delta = \cfail \setminus \cpass$ is paritioned into $n$ disjoint subsets $\Delta_1, ..., \Delta_n$ much as in \ddp.
+Then the follwing cases are evaluated:
+
+\begin{description}
+  	\item[``reduce to subset''] If adding any $\Delta_i$ to the current \cpass\ yields a failing configuration, this configuration can be used as a new \cfail\ as it is smaller than the current \cfail.
+Recurse with a partition granularity of 2.
+  	\item[``increase to complement''] If removing any $\Delta_i$ from the current \cfail\ yields a passing configuration, this configuration can be used as a new \cpass\ as it is larger than the current \cpass.
+Recurse with a partition granularity of 2.
+  	\item[``increase to subset''] If adding any $\Delta_i$ to the current \cpass\ yields a passing configuration, this configuration can be used as a new \cpass\ as it is larger than the current \cpass.
+Recurse with a partition granularity of $n-1$, as the difference changes only by $\Delta_i$ and thus all coarser partitions would result in unresolved tests again.
+  	\item[``reduce to complement''] If removing any $\Delta_i$ from the current \cfail\ yields a failing configuration, this configuration can be used as a new \cfail\, as it is smaller than the current \cfail.
+Recurse with a partition granularity of $n-1$ (s.a.).
+  	\item[``increase granularity''] If none of the above cases apply, i.e. the tests all remain unresolved, then the granularity of the partition of $\Delta$ is increased as in \ddp. 
+  	\item[``done''] If $n \ge |\Delta|$ the granularity can not be increased further.
+In that case every $\Delta_i$ only contains one change.
+If \cpass\ cannot be increased and \cfail\ not be reduced at that point, the difference between them is 1-minimal and they are returned.\\
+\end{description}
+The full \dddiff\ algorithm is shown in Figure \ref{fig:dddiff}.\\
+
+Isolation has two technical advantages over simplification:
+\begin{enumerate}
+	\item It takes less tests to find a 1-minimal failure-inducing difference than to find a minimal failure-inducing subset of changes. %TODO quote paper
+	This is because isolation can improve the current configurations if either a test fails or a test passes whereas simplification needs a test to fail to reduce the size of the current subset of changes.
+Simply put, isolation approaches the solution from both sides simultaneously while simplification only works in one way.
+	\item The simplification algorithm \ddp\ relies on monotone configurations, i.e.
+$test(c)=\cmark \Rightarrow (\forall c' \subseteq c : test(c') = \cmark)$ holds for all $c \in \C$.
+The isolation algorithm \dddiff\ does not need to assume all configurations are monotone.
+With some adaptations it can still make use of that property if it is given, though.
+%TODO quote paper
+\end{enumerate}
 
@@
 In many cases this minimal input would be some sort of "empty" input, be it no data, zero values, etc.
-This minimal input is regarded as the \yd state.
-  	\item The known failing input is regarded as the \td state.
+This minimal input is regarded as the \yd\ state.
+  	\item The known failing input is regarded as the \td\ state.
   	\item A change is defined as the smallest possible difference between any two possible inputs, usually a letter, number or other token of data.
-$\C$ can then be any set of changes that convert the minimal input \yd to the failing input \td.
+$\C$ can then be any set of changes that convert the minimal input \yd\ to the failing input \td.
   	\item The $test$ function returns $\xmark$ if exactly the known error occurs on program execution. It returns $\cmark$ if no error occurs and $\qmark$ in all other cases. 
@@
 
-Consider as an example the C program \reffail which caused the GNU C Compiler (GCC) in version 2.95.2 (on Intel-Linux with optimization enabled) to crash.
+Consider as an example the C program \reffail\ which caused the GNU C Compiler (GCC) in version 2.95.2 (on Intel-Linux with optimization enabled) to crash.
 The program looks like correct C code that should be compiled properly.
-We'd like to use Delta Debugging to find out which part of \reffail is responsible for the crash.
-An empty C program is regarded as the \yd state and \reffail as the \td state.
+We'd like to use Delta Debugging to find out which part of \reffail\ is responsible for the crash.
+An empty C program is regarded as the \yd\ state and \reffail\ as the \td\ state.
 A change is defined as a C language token.
@@
 
-It can be of great help to the software developer to be able to automatically determine the specific parts of the program input that cause the failure.
+It can be of great help for the software developer to be able to automatically determine the specific parts of the program input that cause the failure.
 The information gathered from the above method can be used by the developer to manually trace the responsible input data to the crash.
@@
 Delta Debugging can help trace the error from the root to the failure during program execution as well, though.
-Given a test framework that is able to access program states during execution, DD is used on the program states of two program runs \rpass\ and \rfail\ at significant locations to deduce a \textit{Cause-Effect-Chain (CEC)}.
-
-\definition \defsub{Program states} A program state
+Given a test framework that is able to access \href{def:program_state}{program states} during execution, DD is used on the program states of two \href{def:program_run}{program runs} \rpass\ and \rfail\ at significant locations to deduce a \textit{Cause-Effect-Chain (CEC)}.
 
@@
 
-\definition \defsub{Circumstances and program runs} Let $R$ be the set of all possible circumstance configurations.
-A run is defined as a configuration of circumstances, so every $r \in R$ is called a run.
+\definition \label{def:program_run} \defsub{Circumstances and program runs} Let $R$ be the set of all possible circumstance configurations.
+A run is defined as a configuration of circumstances, so every $r \in R$ is called a run.\\
+
+For our intents and purposes it is sufficient to only consider the program input as the changeable circumstances as we assume the error lies within the program code and occurs independently of operating systems, hardware and other environmental factors.
+
+\definition \label{def:program_state} \defsub{Program states} A program state $p$ is a set of \textit{(variable, value)} pairs that unambiguously determine further program execution.
+The program state always corresponds to a specific point in a \href{def:program_run}{program run}.\\
+
+In order to use Delta Debugging on program states we need to once again define concrete realizations of the abstract DD concepts:
+
+\begin{itemize}
+	\item We assume a program run \rfail\ based on some known faulty input that crashes the program with a certain error.
+We further assume knowledge of a passing program run \rpass\ that runs and terminates correctly.
+	\item We use Delta Debugging to compare program states of \rfail\ and \rpass\ at multiple significant locations \footnote{The number of comparisons vary depending on program size.
+It defines the granularity of the resulting Cause-Effect-Chain.
+The location of a comparison should give an overview of the state of the execution and should be reached by most program runs.
+The locations define the informative power of the Cause Effect Chain.}.
+Each time Delta Debugging is used the program state \sfail\ of \rfail\ is regarded as \td\ and the state \spass\ of \rpass\ is regarded as \yd.
+	\item A change between two program states is defined as either a different value of a variable or a variable that only exists in one program state.
+	\item The set $\C$ contains all such changes between \spass\ and \sfail\ for each comparison.
+	\item The \textit{test} function models further program execution from a given program state.
+It returns \xmark\ if the program crashes with the exact known error and \cmark\ if the program terminates correctly.
+In all other cases \qmark\ is returned (e.g.
+a crash with a different error).
+\end{itemize}
+
+The definition of the test function makes evident why a testing framework is needed.
+We need to be able to extract the program state \spass\ of \rpass\ and \sfail\ of \rfail\ at a certain point in order to compare them.
+More importantly, though, the program code needs to be executed from that point repeatedly with different program states.
+The states might even be artificial, i.e.
+they are created by applying arbitrary subsets of $\C$ to \spass\ and would never occur in any run without outside interference.
+In that case unexpected behaviour will occur and memory leaks are likely.
+The framework needs to make sure no harm is done to the machine or other software.
+Frameworks for testing and analyzing program execution with these capabilities luckily do exist already, though.
+
+Given the above concrete definitions the \href{fig:ddp}{\ddp} algorithm can be used repeatedly to compare \rpass\ and \rfail.
+The result of each execution of \ddp\ will be a set of \textit{(variable, value)} pairs in \sfail\ that are responsible for the failure down the line.
+A Cause-Effect-Chain from the input to the failure is then created by chaining together the results of each state comparison.
+The CEC can give a detailled look into the root cause of the error.\\
+
+Consider as an example the above example program \reffail\ that crashes GCC.
+Applying GCC on $fail.c$ is regarded as \rfail\ and applying GCC to an empty C program as \rpass\ (compiling no code obviously terminates correctly).
+Andreas Zeller's experimental \textit{HOWCOME} algorithm\footnote{The \textit{HOWCOME} algorithm uses Delta Debugging as desribed above and incorporates technical issues like the extraction of program states and applying the test and analysis framework.} %TODO quote
+compares the runs at the function \textit{main} shortly after the start of the program, at the function $combine\_instructions$ in the middle of the program run and at the last hit of the function $if\_then\_else\_cond$ shortly before the failure.
+\textit{HOWCOME} yields a \href{fig:failcec}{Cause-Effect-Chain} that includes the comparisons at those locations.
+
+\begin{figure}[h!]
+	\fbox{
+		\begin{minipage}{\textwidth}
+			\begin{enumerate}
+			\item Execution reaches \textbf{main}. \\Since the program was invoked as ``ccl -O fail.i'' variable \textbf{$argv[2]$} is now \textbf{``fail.i''}
+			\item Execution reaches \textbf{combine\_instructions}.
+\\ Since $argv[2]$ was ``fail.i'', variable \textbf{$*first\_loop\_store\_insn\cecarrow fld[1].rtx \cecarrow fld[1].rtx \cecarrow fld[3].rtx \cecarrow fld[1].rtx$} is now \textbf{$\langle new rtx\_def\rangle$}.
+			\item Execution reaches \textbf{if\_then\_else\_cond (95th hit)}.
+\\ Since $*first\_loop\_store\_insn\cecarrow fld[1].rtx \cecarrow fld[1].rtx \cecarrow fld[3].rtx \cecarrow fld[1].rtx$ was $\langle new rtx\_def\rangle$, variable \textbf{$link \cecarrow fld[0].rtx \cecarrow fld[0].rtx$} is now \textbf{$link$}.
+			\item Execution ends. \\ Since variable $link \cecarrow fld[0].rtx \cecarrow fld[0].rtx$ was $link$, the program now \textbf{terminates with a SIGSEGV signal}. The program fails.
+			\end{enumerate}
+		\end{minipage}
+	}
+	\caption{The Cause-Effect-Chain that is deduced by \textit{HOWCOME} for \reffail.}
+	\label{fig:failcec}
+\end{figure}
+
+The CEC contains the cause for the failure but the fault in the program has to be extracted by the programmer as it is a question of intended vs faulty states.
+The programmer uses the CEC to find the transition between the last intended and the first faulty program state.
+In between these two states a more detailled analysis can be conducted using a finer granularity of compared program states which eventually leads to the error in the program itself.
+The distinction between intended and faulty states may also be conducted by heuristics which would allow the system to automatically search with a finer granularity in the desired section of the runs.
 
 \subsection{Minimizing random unit test cases}
-\subsubsection{The ddmin algorithm}
+
+An important part of software testing today is randomized tests.
+Given a unit under test (UUT) a randomized test case generator will create random sequences of instructions to the UUT with randomized input data.
+The test driver will use the generated sequences to test the functionality of the UUT even under implausible circumstances.
+Non-randomized test drivers with test cases written by the programmer will not be able to cover all these cases (see Fuzzing).%TODO quote paper
+
+The problem with randomized unit testing is that the randomly generated failing test cases are often not useful for the programmer.
+When debugging the code, the failing test cases hold little value as they tend to contain lots of unnecessary instructions, which hides the reason for the failure.
+
+After generating random test cases it would be useful to minimize the failing test cases so they only contain the instructions that are relevant for inducing the failure.
+This is where Delta Debugging comes to play: By viewing a empty test case as the \yd\ state and the randomly generated failing test case as the \td\ state, we can apply a Delta Debugging algorithm to minimize the test case.
+We define the abstract DD concepts in this context more concretely:
+
+\begin{itemize}
+	\item Every instruction in the randomly generated failing test case is regarded as one change.
+	\item $\C$ is regarded as all instructions in the test case in order.
+	\item The \textit{test} function models the execution of the instructions in the test driver.
+The instructions are always executed in the same order as they were in the original randomly generated failing test case.
+\end{itemize}
+
+Both \ddp\ and \dddiff\ can be used in this scenario.
+\ddp\ will return a minimal sequence of instruction that induces the same error as the original test case whereas \dddiff\ returns two test cases, one passing and one failing, with a minimal difference.
+\dddiff\ will likely be more efficient %TODO quote
+ but it remains a matter of context which result is more useful for the debugging of the program code.\\
+
+To improve the runtime of the test case minimization static slicing can be initially applied to the test cases.
+In static slicing Read-Write-Data-Dependencies between the instructions are analyzed and only the instructions that could influence the last instruction remain in the test.
+This method is a static analysis of the test cases, meaning it does not actually need to execute the tests in order to find instructions that are irrelevant for the failure.
+This gives static slicing a runtime advantage compared to minimization by Delta Debugging.
+At the same time static slicing is not as effective as DD minimization.
+A combination of both, that is first applying static slicing and then DD minimization, will retain the advantages of both methods, though: It has a lower runtime but is just as efficient as DD minimization.
+%TODO quote efficient unit test case minimization
 

### 2019-01-13 Paper/article.tex
Added citations.

@@
 
+\ddp\ returns a minimal failure-inducing subset of changes as long as the subset is safe, that is they can be applied to the baseline or removed from \td\ without causing an inconsistency.
+\ddp\ has linear time complexity \cite{Zeller:1999:YMP:318774.318946}.
+
 \subsection{Isolating a minimal difference: \dddiff}
@@
 \end{description}
-The full \dddiff\ algorithm is shown in Figure \ref{fig:dddiff}.\\
+The full \dddiff\ algorithm is shown in Figure \ref{fig:dddiff}.
+It is desribed by Zeller as the successor to \ddp\ and is recommended to generally replace \ddp\ \cite{Zeller:2002:SIF:506201.506206}.
+\\
 
@@
 \begin{enumerate}
-	\item It takes less tests to find a 1-minimal failure-inducing difference than to find a minimal failure-inducing subset of changes. %TODO quote paper
+	\item It takes less tests to find a 1-minimal failure-inducing difference than to find a minimal failure-inducing subset of changes \cite{Zeller:2002:SIF:506201.506206}.
 	This is because isolation can improve the current configurations if either a test fails or a test passes whereas simplification needs a test to fail to reduce the size of the current subset of changes.
@@
 The isolation algorithm \dddiff\ does not need to assume all configurations are monotone.
-With some adaptations it can still make use of that property if it is given, though.
-%TODO quote paper
+With some adaptations it can still make use of that property if it is given, though \cite{Zeller:2002:SIF:506201.506206}.
 \end{enumerate}
@@
 Applying GCC on $fail.c$ is regarded as \rfail\ and applying GCC to an empty C program as \rpass\ (compiling no code obviously terminates correctly).
-Andreas Zeller's experimental \textit{HOWCOME} algorithm\footnote{The \textit{HOWCOME} algorithm uses Delta Debugging as desribed above and incorporates technical issues like the extraction of program states and applying the test and analysis framework.} %TODO quote
+Andreas Zeller's experimental \textit{HOWCOME} algorithm\footnote{The \textit{HOWCOME} algorithm uses Delta Debugging as desribed above and incorporates technical issues like the extraction of program states and applying the test and analysis framework \cite{Zeller:2002:ICC:587051.587053}.}
 compares the runs at the function \textit{main} shortly after the start of the program, at the function $combine\_instructions$ in the middle of the program run and at the last hit of the function $if\_then\_else\_cond$ shortly before the failure.
@@
 	}
-	\caption{The Cause-Effect-Chain that is deduced by \textit{HOWCOME} for \reffail.}
+	\caption{The Cause-Effect-Chain that is deduced by \textit{HOWCOME} for \reffail \cite{Zeller:2002:ICC:587051.587053}.}
 	\label{fig:failcec}
@@
 The test driver will use the generated sequences to test the functionality of the UUT even under implausible circumstances.
-Non-randomized test drivers with test cases written by the programmer will not be able to cover all these cases (see Fuzzing).%TODO quote paper
+Non-randomized test drivers with test cases written by the programmer will not be able to cover all these cases.
 
@@
 After generating random test cases it would be useful to minimize the failing test cases so they only contain the instructions that are relevant for inducing the failure.
-This is where Delta Debugging comes to play: By viewing a empty test case as the \yd\ state and the randomly generated failing test case as the \td\ state, we can apply a Delta Debugging algorithm to minimize the test case.
+This is where Delta Debugging comes to play (\cite{Lei:2005:MRU:1104997.1105255}): By viewing a empty test case as the \yd\ state and the randomly generated failing test case as the \td\ state, we can apply a Delta Debugging algorithm to minimize the test case.
 We define the abstract DD concepts in this context more concretely:
@@
 \ddp\ will return a minimal sequence of instruction that induces the same error as the original test case whereas \dddiff\ returns two test cases, one passing and one failing, with a minimal difference.
-\dddiff\ will likely be more efficient %TODO quote
+\dddiff\ will likely be more efficient (\cite{Zeller:2002:SIF:506201.506206})
  but it remains a matter of context which result is more useful for the debugging of the program code.\\
@@
 At the same time static slicing is not as effective as DD minimization.
-A combination of both, that is first applying static slicing and then DD minimization, will retain the advantages of both methods, though: It has a lower runtime but is just as efficient as DD minimization.
-%TODO quote efficient unit test case minimization
+A combination of both, that is first applying static slicing and then DD minimization, will retain the advantages of both methods, though: It has a lower runtime but is just as efficient as DD minimization \cite{Leitner:2007:EUT:1321631.1321698}.
 
@@
 
+\bibliography{article}
+
 \end{document}

### 2019-01-14 Paper/article.tex
Added conclusion and improved smaller aspects like links.

@@
 % fail.c reference shortcut
-\newcommand{\reffail}[0]{\href{fig:fail.c}{\ensuremath{fail.c^{Fig. \ref{fig:fail.c}}}}}
+\newcommand{\reffail}[0]{\hyperref[fig:fail.c]{\ensuremath{fail.c}}}
 
@@
 \begin{abstract}
-	This paper deals with the Delta Debugging method.
+	Debugging is tedious work for any programmer.
+This thought sparked the concept of Delta Debugging.
+In the late 90's and early 2000's Andreas Zeller abstracted a simplistic scenario (``\yd\ my program worked, \td\ it does not.
+Why?'') into algorithmically tackling debugging.
+The goal is simple: If every change between \yd\ and \td\ is seen as a Delta, which Deltas are responsible for the failure of \td?
+
+	This paper summarizes Zeller's early work on Delta Debugging \cite{Zeller:1999:YMP:318774.318946} and further gives an overview about several use cases for Delta Debugging that have been found in the early 2000's.
 \end{abstract}
@@
 \section{What is Delta Debugging?}
+\label{sec:what_is_dd}
+
+This section outlines the scenario that acts as the base for all following definitions and gives an idea of how Delta Debugging is useful.
 
@@
 
-\section{Definitions and basework}
+\section{Formalizing concepts}
+\label{sec:formalizing_concepts}
 
@@
 \section{Incrementally approaching a plausible DD algorithm}
+\label{sec:ddalgorithms}
 
@@
 A na\"ive Delta Debugging algorithm called \sdd\ will be refined into the \dd\ algorithm and the \ddp\ algorithm in this chapter in order to describe and solve both difficulties.
-Finally, isolation of a minimal-failure inducing difference, an alternative approach to DD, is protrayed through the \dddiff\ algorithm.
+Finally, isolation of a minimal-failure inducing difference, an alternative approach to DD, is portrayed through the \dddiff\ algorithm.
 
@@
 It becomes evident that this algorithm is not effective because it is only capable of finding single changes that cause failure on their own.
-But what if mutiple changes pass the test individually but only their combination induces failure?
+But what if multiple changes pass the test individually but only their combination induces failure?
 \\
@@
 
-\ddp\ returns a minimal failure-inducing subset of changes as long as the subset is safe, that is they can be applied to the baseline or removed from \td\ without causing an inconsistency.
-\ddp\ has linear time complexity \cite{Zeller:1999:YMP:318774.318946}.
+\lemma \ddp\ returns a minimal failure-inducing subset of changes as long as the subset is safe, that is they can be applied to the baseline or removed from \td\ without causing an inconsistency.
+\ddp\ has linear time complexity.
+
+\proof See \cite{Zeller:1999:YMP:318774.318946}.
 
@@
 The algorithm is initialized with $\cpass = \emptyset$ and $\cfail = \C$.
-The difference $\Delta = \cfail \setminus \cpass$ is paritioned into $n$ disjoint subsets $\Delta_1, ..., \Delta_n$ much as in \ddp.
-Then the follwing cases are evaluated:
+The difference $\Delta = \cfail \setminus \cpass$ is partitioned into $n$ disjoint subsets $\Delta_1, ..., \Delta_n$ much as in \ddp.
+Then the following cases are evaluated:
 
@@
 The full \dddiff\ algorithm is shown in Figure \ref{fig:dddiff}.
-It is desribed by Zeller as the successor to \ddp\ and is recommended to generally replace \ddp\ \cite{Zeller:2002:SIF:506201.506206}.
+It is described by Zeller as the successor to \hyperref[fig:ddp]{\ddp} and is recommended to generally replace \ddp\ \cite{Zeller:2002:SIF:506201.506206}.
 \\
@@
 \section{Exemplary use cases}
+\label{sec:use_cases}
 
@@
 The information gathered from the above method can be used by the developer to manually trace the responsible input data to the crash.
-Considering how a miniscule input change can cause entirely different execution paths this task is tedious if not impossible.
+Considering how a minuscule input change can cause entirely different execution paths this task is tedious if not impossible.
 Delta Debugging can help trace the error from the root to the failure during program execution as well, though.
-Given a test framework that is able to access \href{def:program_state}{program states} during execution, DD is used on the program states of two \href{def:program_run}{program runs} \rpass\ and \rfail\ at significant locations to deduce a \textit{Cause-Effect-Chain (CEC)}.
+Given a test framework that is able to access \hyperref[def:program_state]{program states} during execution, DD is used on the program states of two \hyperref[def:program_run]{program runs} \rpass\ and \rfail\ at significant locations to deduce a \textit{Cause-Effect-Chain (CEC)}.
 
@@
 \definition \label{def:program_state} \defsub{Program states} A program state $p$ is a set of \textit{(variable, value)} pairs that unambiguously determine further program execution.
-The program state always corresponds to a specific point in a \href{def:program_run}{program run}.\\
+The program state always corresponds to a specific point in a \hyperref[def:program_run]{program run}.\\
 
@@
 they are created by applying arbitrary subsets of $\C$ to \spass\ and would never occur in any run without outside interference.
-In that case unexpected behaviour will occur and memory leaks are likely.
+In that case unexpected behavior will occur and memory leaks are likely.
 The framework needs to make sure no harm is done to the machine or other software.
@@
 
-Given the above concrete definitions the \href{fig:ddp}{\ddp} algorithm can be used repeatedly to compare \rpass\ and \rfail.
+Given the above concrete definitions the \hyperref[fig:ddp]{\ddp} algorithm can be used repeatedly to compare \rpass\ and \rfail.
 The result of each execution of \ddp\ will be a set of \textit{(variable, value)} pairs in \sfail\ that are responsible for the failure down the line.
 A Cause-Effect-Chain from the input to the failure is then created by chaining together the results of each state comparison.
-The CEC can give a detailled look into the root cause of the error.\\
+The CEC can give a detailed look into the root cause of the error.\\
 
@@
 Applying GCC on $fail.c$ is regarded as \rfail\ and applying GCC to an empty C program as \rpass\ (compiling no code obviously terminates correctly).
-Andreas Zeller's experimental \textit{HOWCOME} algorithm\footnote{The \textit{HOWCOME} algorithm uses Delta Debugging as desribed above and incorporates technical issues like the extraction of program states and applying the test and analysis framework \cite{Zeller:2002:ICC:587051.587053}.}
+Andreas Zeller's experimental \textit{HOWCOME} algorithm\footnote{The \textit{HOWCOME} algorithm uses Delta Debugging as described above and incorporates technical issues like the extraction of program states and applying the test and analysis framework \cite{Zeller:2002:ICC:587051.587053}.}
 compares the runs at the function \textit{main} shortly after the start of the program, at the function $combine\_instructions$ in the middle of the program run and at the last hit of the function $if\_then\_else\_cond$ shortly before the failure.
-\textit{HOWCOME} yields a \href{fig:failcec}{Cause-Effect-Chain} that includes the comparisons at those locations.
+\textit{HOWCOME} yields a \hyperref[fig:failcec]{Cause-Effect-Chain} that includes the comparisons at those locations.
 
@@
 	}
-	\caption{The Cause-Effect-Chain that is deduced by \textit{HOWCOME} for \reffail \cite{Zeller:2002:ICC:587051.587053}.}
+	\caption{The Cause-Effect-Chain that is deduced by \textit{HOWCOME} for \reffail\ \cite{Zeller:2002:ICC:587051.587053}.}
 	\label{fig:failcec}
@@
 The programmer uses the CEC to find the transition between the last intended and the first faulty program state.
-In between these two states a more detailled analysis can be conducted using a finer granularity of compared program states which eventually leads to the error in the program itself.
+In between these two states a more detailed analysis can be conducted using a finer granularity of compared program states which eventually leads to the error in the program itself.
 The distinction between intended and faulty states may also be conducted by heuristics which would allow the system to automatically search with a finer granularity in the desired section of the runs.
@@
 
-Both \ddp\ and \dddiff\ can be used in this scenario.
+Both \hyperref[fig:ddp]{\ddp} and \hyperref[fig:dddiff]{\dddiff} can be used in this scenario.
 \ddp\ will return a minimal sequence of instruction that induces the same error as the original test case whereas \dddiff\ returns two test cases, one passing and one failing, with a minimal difference.
@@
 
-To improve the runtime of the test case minimization static slicing can be initially applied to the test cases.
+Static slicing can be initially applied to the test cases to improve the runtime of the test case minimization.
 In static slicing Read-Write-Data-Dependencies between the instructions are analyzed and only the instructions that could influence the last instruction remain in the test.
@@
 
-\section{Reflection and conclusion}
-\subsection{Abstractness: strength and weakness}
-\subsection{Conclusion}
+\section{Contribution}
+\label{sec:contribution}
+
+This paper is not a research paper.
+It only summarizes work of others from the given citations and does not incorporate any research conducted by the author.
+The content of sections \ref{sec:what_is_dd}, \ref{sec:formalizing_concepts} and \ref{sec:ddalgorithms} is mostly taken from \cite{Zeller:1999:YMP:318774.318946}.
+\dddiff\ is taken from \cite{Zeller:2002:SIF:506201.506206} where it is called \textit{dd}.
+The use cases in section \ref{sec:use_cases} and most importantly the $fail.c$ case studies have been taken from \cite{Zeller:2002:ICC:587051.587053} and \cite{Lei:2005:MRU:1104997.1105255}.
+Static slicing in section \ref{sec:use_cases} is a concept portrayed in \cite{Leitner:2007:EUT:1321631.1321698}.
+
+\section{Conclusion}
+\label{sec:conclusion}
+Zeller initially defined Delta Debugging as abstract as possible \cite{Zeller:1999:YMP:318774.318946}.
+This allows the method to be applied to a variety of use cases in testing and debugging as outlined in section \ref{sec:use_cases}.
+The abstractness is also a weakness, though, as in praxis every time Delta Debugging is applied a variety of these abstract concepts need to be concretized.
+Bringing the Delta Debugging algorithms that are portrayed in section \ref{sec:ddalgorithms} into a practical context might still be a lot of work to the human debugger.
+
+Nonetheless Delta Debugging has its rightful place in the scientific approach to debugging as the idea continues to find new uses today, almost 20 years after its introduction by Zeller (The \href{https://dl.acm.org/dl.cfm}{ACM Digital Library} mentions 6 papers directly incorporating Delta Debugging in 2017/2018).
+An important development of Delta Debugging is Hierarchical Delta Debugging \cite{Misherghi:2006:HHD:1134285.1134307} which speeds up the process by making use of hierarchical structures in data, e.g.
+when debugging XML code.
 

### 2019-02-06 Vortrag/slides/sections/cause-effect-chains.tex
More notes and nicer use case sections.
Minor changes.

@@
-\subsection{The DD principle as a tool}
-\begin{frame}{Using DD to find the reason a program fails}
+\subsection{Use case 2: Constructing Cause-Effect-Chains}
+\begin{frame}{Use case 2: Finding the reason a program fails}
 	\begin{itemize}
@@
 
-\begin{frame}{Using DD to find the reason a program fails}
+\begin{frame}{Use case 2: Finding the reason a program fails}
 	\begin{alertblock}{Difficulty: Causalities from the input change to the failure}
@@
 
+	\bigskip
 	\pause
@@
 		\end{itemize}
-		\item With this the runs $r_\cmark$ and $r_\xmark$ can be compared in different significant locations. 
+		%\item With this the runs $r_\cmark$ and $r_\xmark$ can be compared in different significant locations. 
 		\item We use delta debugging to find the current $(variable, value)$ pairs that are responsible for the failure of $r_\xmark$ further down the line.
@@
 			\onslide<2->{\begin{enumerate}
-				\only<2>{\item A \red{cycle in the RTL tree} of the GCC compiler is created at the 95th hit of \textbf{if\_then\_else\_cond}.}
-				\onslide<3>{\item A cycle in the RTL tree of the GCC compiler is created at the 95th hit of \textbf{if\_then\_else\_cond}.} 
+				\only<2>{\item A \red{cycle in the RTL tree} of GCC is created at the 95th hit of \textbf{if\_then\_else\_cond}.}
+				\onslide<3>{\item A cycle in the RTL tree of GCC is created at the 95th hit of \textbf{if\_then\_else\_cond}.} 
 				\onslide<3>{\item The \red{cycle produces the error}.}

### 2019-02-08 Vortrag/slides/sections/cause-effect-chains.tex

@@
 \subsection{Use case 2: Constructing Cause-Effect-Chains}
-\begin{frame}{Use case 2: Finding the reason a program fails}
-	\begin{itemize}
-		\item The \ddp\ algorithm as described above can be used to find out which part of a certain input is responsible for crashing a program.
-		\item From a developers perspective it would be great to know \textit{why} this specific input induces failure.
-	\end{itemize}
+%\begin{frame}{Use case 2: Finding the reason a program fails}
+%	\begin{itemize}
+%		\item The \ddp\ algorithm as described above can be used to find out which part of a certain input is responsible for crashing a program.
+%		\item From a developers perspective it would be great to know \textit{why} this specific input induces failure.
+%	\end{itemize}
 
-\end{frame}
+%\end{frame}
 
 \begin{frame}{Use case 2: Finding the reason a program fails}
+
 	\begin{alertblock}{Difficulty: Causalities from the input change to the failure}
@@
 
+	\pause
 	\bigskip
-	\pause
 
@@
 	\end{exampleblock}
+
 \end{frame}
@@
 			\item Execution reaches \textbf{if\_then\_else\_cond (95th hit)}.
-\\ Since $*first\_loop\_store\_insn\cecarrow fld[1].rtx \cecarrow fld[1].rtx \cecarrow fld[3].rtx \cecarrow fld[1].rtx$ was $\langle new rtx\_def\rangle$, variable \only<3>{\textbf{$link \cecarrow fld[0].rtx \cecarrow fld[0].rtx$} is now \textbf{$link$}} \only<2>{\red{\textbf{$link \cecarrow fld[0].rtx \cecarrow fld[0].rtx$} is now \textbf{$link$}}}.
+\\ Since $*first\_loop\_store\_insn\cecarrow fld[1].rtx \cecarrow fld[1].rtx \cecarrow fld[3].rtx \cecarrow fld[1].rtx$ was $\langle new rtx\_def\rangle$, variable \only<1,3>{\textbf{$link \cecarrow fld[0].rtx \cecarrow fld[0].rtx$} is now \textbf{$link$}} \only<2>{\red{\textbf{$link \cecarrow fld[0].rtx \cecarrow fld[0].rtx$} is now \textbf{$link$}}}.
 			\item Execution ends.

### 2019-03-21 Paper/article.tex
Incorporated Stephans comments and final tweaks.

@@
 
-\author{Moritz C. Laupichler}{Fakultät für Informatik, Karlsruhe Institute of Technology, Germany}{moritz.laupichler@student.kit.edu}{}{}
+\author{Moritz C. Laupichler}{Department of Informatics, Karlsruhe Institute of Technology, Germany}{moritz.laupichler@student.kit.edu}{}{}
 
@@
 \usepackage{xcolor}
-\definecolor{td-green}{rgb}{0,.59,.51}
+\definecolor{td-green}{rgb}{0,.5, 0}
 
@@
 	Debugging is tedious work for any programmer.
-This thought sparked the concept of Delta Debugging.
-In the late 90's and early 2000's Andreas Zeller abstracted a simplistic scenario (``\yd\ my program worked, \td\ it does not.
-Why?'') into algorithmically tackling debugging.
-The goal is simple: If every change between \yd\ and \td\ is seen as a Delta, which Deltas are responsible for the failure of \td?
+This thought sparked the concept of Delta Debugging which was first described by Andreas Zeller in the late 90's and early 2000's.
+A simplistic scenario (``\yd\ my program worked, \td\ it does not.
+Why?'') was abstracted into algorithmically tackling debugging.
+The goal is simple: If every change between \yd\ and \td\ is seen as a Delta, which Deltas are responsible for the failure \td?
 
@@
 
-This section outlines the scenario that acts as the base for all following definitions and gives an idea of how Delta Debugging is useful.
+This section outlines the scenario that acts as the base for all following definitions and gives an idea of the goal of Delta Debugging.
 
@@
 Imagine a software development process where the latest changes to the code cause known tests to fail even though an older version of the code passed these exact tests.
-As the basis of the Delta Debugging principle we need to define a simple abstraction of this point in the development process: We know a state \yd\ that passes the tests and a state \td\ that fails the tests.
-Furthermore we know all changes that when applied to \yd\ yield \td.
-\\
+As the base of the Delta Debugging principle we need to define a simple abstraction of this point in the development process: We know a state \yd\ that passes the tests and a state \td\ that fails the tests.
+Furthermore we know all changes that yield \td\ when applied to \yd.\\
 The goal of Delta Debugging is to automate extraction of the exact changes to \yd\ that cause the failure in \td.
@@
 \definition \defsub{Failure-inducing configuration} A configuration $c$ is called \textbf{failure-inducing} iff 
-\[ \forall c' (c \subseteq c' \subseteq \C \rightarrow test(c') \ne \cmark) \] 
+\[ \forall c' : c \subseteq c' \subseteq \C \Rightarrow test(c') \ne \cmark \] 
 holds.
@@
 \definition \defsub{Minimal failure-inducing configuration} A failure-inducing configuration $c$ is \textbf{minimal} iff 
-\[ \forall c' \subset c: test(c') \ne \xmark \]
+\[ \forall c' \subsetneq c: test(c') \ne \xmark \]
 holds.\\
@@
 
-There is two major difficulties when approaching a Delta Debugging algorithm: interference and inconsistency.
+There are two major difficulties when approaching a Delta Debugging algorithm: interference and inconsistency.
 A na\"ive Delta Debugging algorithm called \sdd\ will be refined into the \dd\ algorithm and the \ddp\ algorithm in this chapter in order to describe and solve both difficulties.
@@
 
-The na\"ive algorithm \sdd\ recursively splits \C\ into smaller and smaller parts until one failure-inducing change is found.
-In each recursion step the current set of changes, $c$, is partitioned in two halves, $c_1$ and $c_2$, and both halves are tested.
+The na\"ive algorithm \hyperref[fig:sdd]{\sdd} as shown in Figure \ref{fig:sdd} recursively splits \C\ into smaller and smaller parts until one failure-inducing change is found.
+In each recursion step the current set of changes, $c$, is partitioned into two halves, $c_1$ and $c_2$, and both halves are tested.
 The half that fails the tests is regarded in the next recursion step.
@@
 \fbox{\begin{minipage}{\textwidth}
+$simpledd(\C) = dd_1(\C)$ where \\
 
 \begin{algorithmic}[1]
-		\Function{simpledd}{$c: 2^{\C}$}
+		\Function{$dd_1$}{$c: 2^{\C}$}
 			\If{$|c| = 1$} \Return $c$ \EndIf
-			\State Partition $c$ into two halves $c_1$, $c_2$ so that $c_1 \cap c_2 = \emptyset$
-			\If{($test(c_1) = \xmark$)} \Return $simpledd(c_1)$ \Comment("Found in $c_1$") \Else{} \Return $simpledd(c_2)$ \Comment("Found in $c_2$")
-			\EndIf
+			\State Partition $c$ arbitrarily into two parts $c_1$, $c_2$ such that $|c_1| \approx |c_2| \approx |c|/2$
+			\State \Return $\left \{ \text{\begin{tabular}{l l r}
+				$dd_1(c_1)$ & if $test(c_1) = \xmark$ & \Comment("Found in $c_1$") \\
+				$dd_1(c_2)$ & otherwise & \Comment("Found in $c_2$")
+			\end{tabular}}\right.$
 		\EndFunction
@@
 \\
+
+\begin{samepage}
+
 It becomes evident that this algorithm is not effective because it is only capable of finding single changes that cause failure on their own.
@@
 \definition \defsub{Interference} Two configurations $c_1$ and $c_2$ \textbf{interfere} iff $test(c_1) = \cmark$, $test(c_2) = \cmark$ but $test(c_1 \cup c_2) = \xmark$. \\
+\end{samepage}
 
@@
 
-\sdd\ is unable to handle the case of both $c_1$ and $c_2$ passing.
-A new algorithm \dd\ is created which behaves like \sdd\ but incorporates the case "Interference".
+\sdd\ is unable to handle the case of both $c_1$ and $c_2$ passing i.e.
+interfering configurations.
+A new algorithm \dd\ is created which behaves like \sdd\ but incorporates the case "Interference" as shown in Figure \ref{fig:dd}.
 In this case the algorithm searches for the failure-inducing change in $c_1$ while keeping $c_2$ applied and vice versa.
@@
 
-$dd = dd_2(\C, \emptyset)$ where \\
+$dd(\C) = dd_2(\C, \emptyset)$ where \\
 
@@
 	\Function{$dd_2$}{$c,r:2^\C$} $: 2^{\C}$
-	\State let $c_1,c_2 \subseteq c \text{ with } c_1 \cup c_2 = c, c_1 \cap c_2 = \emptyset, |c_1| \approx |c_2|$
+	\State Partition $c$ arbitrarily into two parts $c_1$, $c_2$ such that $|c_1| \approx |c_2| \approx |c|/2$
 	\State $d_1 = dd_2(c_1, c_2 \cup r)$ \Comment("Search in $c_1$ while leaving $c_2$ applied")
 	\State $d_2 = dd_2(c_2, c_1 \cup r)$ \Comment("Search in $c_2$ while leaving $c_1$ applied")\medskip  
-	\State \Return $ \begin{cases}
-			c & \text{if } |c|=1, \\
-			dd_2(c_1,r) & \text{if } test(c_1 \cup r) = \xmark, \text{\Comment("Found in $c_1$")}\\
-			dd_2(c_2,r) & \text{if } test(c_2 \cup r) = \xmark, \text{\Comment("Found in $c_2$")}\\
-			d_1 \cup d_2 & \text{otherwise } \text{\Comment("Interference")}
-		\end{cases}$
+	\State \Return $\left \{ \text{\begin{tabular}{l l r}
+			$c$ & if $|c|=1$ & \Comment("Found") \\
+			$dd_2(c_1,r)$ & if $test(c_1 \cup r) = \xmark$ & \Comment("Found in $c_1$")\\
+			$dd_2(c_2,r)$ & if $test(c_2 \cup r) = \xmark$ & \Comment("Found in $c_2$")\\
+			$d_1 \cup d_2$ & otherwise & \Comment("Interference")
+		\end{tabular}}
+			\right.$
 	\EndFunction
@@
 
-In the real world of code and regression tests this scenario is most likely caused by a combination of changes resulting in syntactical or semantical errors.
-An inconsistent configuration may also occur when a change is dependent on another change not included in the configuration or when the program execution fails during testing.
+In the real world of code and regression tests this scenario is most likely caused by a combination of changes resulting in syntactical or semantic errors.
+An inconsistent configuration may also occur when a change is dependent on another change, that is not included in the configuration, or when the program execution fails during testing.
 
-\dd\ does not deal with inconsistent configurations. It behaves wrong if the configurations that are tested in the "Found" cases are inconsistent.
+\dd\ does not deal with inconsistent configurations. It delivers wrong results if the configurations that are tested in the "Found" cases are inconsistent.
 
@@
 By applying less changes to \yd\ or removing less changes from \td\ (s.
-case "Preference" below).
-the chances of the result being consistent is increased.
-If the algorithm runs into only inconsistent configurations the number of subsets is increased and the chances of at least one configuration being consistent after is increased.
-This solution requires the final Delta Debugging algorithm to handle any number $n$ of subsets whereas \dd\ only handles two change sets, $c_1$ and $c_2$ at a time.
+case "Preference" below) the chances of the result being consistent are increased.
+If the algorithm runs into only inconsistent configurations the number of subsets is increased, thus increasing the chances of at least one configuration being consistent after.
+This solution requires the new Delta Debugging algorithm to handle any number $n$ of subsets whereas \dd\ only handles two change sets, $c_1$ and $c_2$, at a time.
 
-\definition \defsub{Complement of configuration} Let $c \subseteq \C$ be a configuration.
-The configuration $\bar{c}$ is called the \textbf{complement} of $c$.
-It contains all changes that are contained in the current search but not in $c$.\\
+\definition \defsub{Complement of configuration} Let $c \subseteq \C$ be a configuration that is currently being searched for failure-inducing changes.
+Let $r \subseteq \C$ be the changes that remain applied during the search.
+For a subset $a \subseteq c$ the \textbf{complement} is $\bar{a} = c \setminus (a \cup r)$.\\
 
-The final Delta Debugging algorithm \ddp\ on the subsets $c_1, \dots, c_n$ knows four cases:\\
+The new Delta Debugging algorithm \ddp\ on the subsets $c_1, \dots, c_n$ as shown in Figure \ref{fig:ddp} knows four main cases:\\
 \begin{description}
@@
 Continue search in $c_i$ while leaving $\bar{c_i}$ applied to promote consistency.
-	\item["Try Again"] If none of the above cases apply then continue the search with $2n$ subsets. The chance of consistent configurations is increased.
+	\item["Try Again"] If none of the above cases apply then continue the search with $2n$ subsets. The chances of consistent configurations are increased.
 \end{description}
 
-\begin{figure}[h!]
+\begin{figure}[H]
 \fbox{\begin{minipage}{\textwidth}
 
-$dd^+(c) = dd_3(c, \emptyset,2)$ where \\
+$dd^+(\C) = dd_3(\C, \emptyset,2)$ where \\
 
@@
 	\Function{$dd_3$}{$c,r:2^\C, n: \nN$} $: 2^{\C}$
-	\State let $c_1,\dots,c_n \subseteq c \text{ with } \bigcup c_i = c, c_i \text{ pairwise disjoint}, \forall i: |c_i| \approx |c|/n$
+	\State Partition $c$ arbitrarily into $n$ parts $c_1, \dots, c_n$ such that $\forall i: |c_i| \approx |c|/n$
 	\State let $\bar{c_i} = c \setminus (c_i \cup r), t_i = test(c_i \cup r), \bar{t_i} = test(\bar{c_i} \cup r)$
-	\State let $c' = c \cap \bigcap \set{\bar{c_i} \mid \bar{t_i} = \xmark}, r' = r \cup \bigcup \set{c_i \mid t_i = \cmark}, n' = min (\mid c' \mid, 2n)$
+	\State let $c' = c \cap \bigcap \set{\bar{c_i} \mid \bar{t_i} = \xmark}, r' = r \cup \bigcup \set{c_i \mid t_i = \cmark}, n' = min (|c'|, 2n)$
 	\State let $d_i = dd_3(c_i, \bar{c_i} \cup r, 2), \bar{d_i} = dd_3(\bar{c_i},c_i \cup r, 2)$
-	\State \Return $ \begin{cases}
-			c & \text{if } |c|=1, \text{\Comment("Found")}\\
-			dd_3(c_i, r, 2) & \text{if } t_i = \xmark \text{ for some i \Comment("Found in $c_i$")}\\
-			d_i \cup \bar{d_i} & \text{if } t_i = \cmark \wedge \bar{t_i} = \cmark \text{ for some i \Comment("Interference")}\\
-			d_i & \text{if } t_i = \qmark \wedge \bar{t_i} = \cmark \text{ for some i \Comment("Preference")}\\
-			dd_3(c',r',n') & \text{if } n < \mid c \mid \text{\Comment("Try Again")}\\
-			c' & \text{otherwise \Comment("Nothing Left")}
-		\end{cases}$
+	\State \Return $\left \{ \text{\begin{tabular}{l l r}
+		$c$ & if |c|=1 & \Comment("Found")\\
+		$dd_3(c_i, r, 2)$ & if $t_i = \xmark$ for some $i$ & \Comment("Found in $c_i$")\\
+		$d_i \cup \bar{d_i}$ & if $t_i = \cmark \wedge \bar{t_i} = \cmark$ for some $i$ & \Comment("Interference")\\
+		$d_i$ & if $t_i = \qmark \wedge \bar{t_i} = \cmark$ for some $i$ & \Comment("Preference")\\
+		$dd_3(c',r',n')$ & if $n < |c|$ & \Comment("Try Again")\\
+		$c'$ & otherwise & \Comment("Nothing Left")
+	\end{tabular}}\right.$
 	\EndFunction
@@
 
-\lemma \ddp\ returns a minimal failure-inducing subset of changes as long as the subset is safe, that is they can be applied to the baseline or removed from \td\ without causing an inconsistency.
+\lemma \ddp\ returns a minimal failure-inducing subset of changes as long as the subset is safe, i.e.
+they can be applied to \yd\ or removed from \td\ without causing an inconsistency.
 \ddp\ has linear time complexity.
@@
 Another goal that can be achieved by Delta Debugging is the isolation of the minimal failure-inducing difference between \yd\ and \td.
-Isolation means to replicate this difference through two other configurations \cpass\ and \cfail\ that will not not make any sense from a production standpoint but portray the difference in a minimal and thus simple to understand fashion.
+Isolation means to replicate this difference through two other configurations \cpass\ and \cfail\ that will not make any sense from a production standpoint but portray the difference in a minimal and thus simple to understand fashion.
 This is exactly what the programmer wants when debugging the code: A simplistic reconstruction of what separates passing code from the known failing code in the given context, i.e.
@@
 
-In more technical terms the goal is to find two configurations \cpass and \cfail, such that $\emptyset \subseteq \cpass \subset \cfail \subseteq \C$, $test(\cpass) = \cmark$ and $test(\cfail) = \xmark$ hold and the difference between the configurations $\Delta = \cfail \setminus \cpass$ is minimal.
+In more technical terms the goal is to find two configurations \cpass\ and \cfail, such that $\emptyset \subseteq \cpass \subsetneq \cfail \subseteq \C$, $test(\cpass) = \cmark$ and $test(\cfail) = \xmark$ hold and the difference between the configurations $\Delta = \cfail \setminus \cpass$ is minimal.
 
-\definition \label{def:minimal_difference} \defsub{Minimal failure-inducing difference} Let \cpass\ and \cfail\ be two test cases with $\emptyset \subseteq \cpass \subset \cfail \subseteq \C$.
-Their difference $\Delta = \cfail \setminus \cpass$ is \textbf{minimal} iff \[ \forall \Delta_i \subset \Delta : test(\cpass \cup \Delta_i) \ne \cmark \wedge test(\cfail \setminus \Delta_i) \ne \xmark \]
+\definition \label{def:minimal_difference} \defsub{Minimal failure-inducing difference} Let \cpass\ and \cfail\ be two test cases with $\emptyset \subseteq \cpass \subsetneq \cfail \subseteq \C$.
+Their difference $\Delta = \cfail \setminus \cpass$ is \textbf{minimal} iff \[ \forall \Delta' \subsetneq \Delta : test(\cpass \cup \Delta') \ne \cmark \text{ and } test(\cfail \setminus \Delta') \ne \xmark \]
 
@@
 \definition \label{def:n-minimal_difference} \defsub{N-minimal failure-inducing difference} Let \cpass\ and \cfail\ be defined as above.
-Their difference $\Delta = \cfail \setminus \cpass$ is \textbf{n-minimal} iff \[ \forall \Delta_i \subset \Delta : |\Delta_i| \le n \Rightarrow test(\cpass \cup \Delta_i) \ne \cmark \wedge test(\cfail \setminus \Delta_i) \ne \xmark \]
+Their difference $\Delta = \cfail \setminus \cpass$ is \textbf{n-minimal} iff \[ \forall \Delta' \subsetneq \Delta : |\Delta'| \le n \Rightarrow (test(\cpass \cup \Delta') \ne \cmark \text{ and } test(\cfail \setminus \Delta') \ne \xmark) \]
 
 We aim at finding two configurations with a \textit{1-minimal failure-inducing difference} - that is a configuration \cfail\ where taking away any change $\delta \in \Delta$ will make the failure disappear.\\
-
-\begin{figure}[t]
-	\fbox{
-		\begin{minipage}{\textwidth}
-		The goal of \dddiff\ is to find $(\cfail, \cpass) = \dddiff(\C)$ such that $\emptyset \subseteq \cpass \subset \cfail \subseteq \C$, $test(\cpass) = \cmark$, $test(\cfail) = \xmark$, and $\Delta = \cfail \setminus \cpass$ is 1-minimal.\\
-
-		The \dddiff\ algorithm is $\dddiff(\C)=dd_4(\emptyset, \C, 2)$ where
-
-				\begin{algorithmic}[1]
-					\Function{$dd_4$}{$\cpass,\cfail:2^{\C}$, $n: \nN$}: $2^{\C} \times 2^{\C}$
-					\State let $\Delta = \cfail \setminus \cpass$
-					\State let $\Delta = \bigcup^{n}_{i=1} \Delta_i$ with all $\Delta_i$ pairwise disjoint and $\forall \Delta_i: |\Delta_i| \approx |\Delta|/n$
-					\State let $t_{\cmark, i} = test(\cpass \cup \Delta_i)$ for $i \in \set{1, \dots, n}$ \Comment(``test for larger passing configuration'')
-					\State let $t_{\xmark, i} = test(\cfail \setminus \Delta_i)$ for $i \in \set{1, \dots, n}$ \Comment(``test for smaller failing configuration'')
-					\State \Return $\begin{cases}
-						dd_4(\cpass, \cpass \cup \Delta_i,2) & \text{if } \; t_{\cmark, i} = \xmark \text{\Comment(``reduce to subset'')} \\
-						dd_4(\cfail \setminus \Delta_i, \cfail, 2) & \text{if } t_{\xmark, i} = \cmark \text{\Comment(``increase to complement'')}\\
-						dd_4(\cpass \cup \Delta_i, \cfail, max(n-1, 2)) & \text{if } t_{\cmark, i} = \cmark \text{\Comment(``increase to subset'')}\\
-						dd_4(\cpass, \cfail \setminus \Delta_i, max(n-1,2)) & \text{if } t_{\xmark, i} = \xmark \text{\Comment(``reduce to complement'')}\\
-						dd_4(\cpass, \cfail, min(2n, |\Delta|)) & \text{if } n < |\Delta| \text{\Comment(``increase granularity'')} \\
-						(\cpass, \cfail) & \text{otherwise \Comment(``done'')} 
-					\end{cases}$
-					(check the first cases for all $i \in \set{1, \dots, n}$)
-					\EndFunction
-				\end{algorithmic}
-		\end{minipage}
-	}
-	\caption{The Delta Debugging algorithm for finding two configurations with minimal difference, \dddiff.}
-	\label{fig:dddiff}
-\end{figure}
 
@@
 The full \dddiff\ algorithm is shown in Figure \ref{fig:dddiff}.
-It is described by Zeller as the successor to \hyperref[fig:ddp]{\ddp} and is recommended to generally replace \ddp\ \cite{Zeller:2002:SIF:506201.506206}.
+It is described by Zeller as the successor to \hyperref[fig:ddp]{\ddp} and is recommended to generally replace \ddp\ as it is superior in performance and result quality \cite{Zeller:2002:SIF:506201.506206}.
 \\
+
+\begin{figure}[H]
+	\fbox{
+		\begin{minipage}{\textwidth}
+
+		$\dddiff(\C)=dd_4(\emptyset, \C, 2)$ where \\
+
+				\begin{algorithmic}[1]
+					\Function{$dd_4$}{$\cpass,\cfail:2^{\C}$, $n: \nN$}: $2^{\C} \times 2^{\C}$
+					\State let $\Delta = \cfail \setminus \cpass$
+					\State Partition $\Delta$ arbitrarily into $n$ parts $\Delta_1, \dots, \Delta_n$ such that $\forall i: |\Delta_i| \approx |\Delta|/n$
+					\State let $t_{\cmark, i} = test(\cpass \cup \Delta_i)$ for $i \in \set{1, \dots, n}$ \Comment(``test for larger passing configuration'')
+					\State let $t_{\xmark, i} = test(\cfail \setminus \Delta_i)$ for $i \in \set{1, \dots, n}$ \Comment(``test for smaller failing configuration'')
+					\State \Return $\left \{ \text{\begin{tabular}{l l r}
+						$dd_4(\cpass, \cpass \cup \Delta_i,2)$ & if $t_{\cmark, i} = \xmark$ & \Comment(``reduce to subset'') \\
+						$dd_4(\cfail \setminus \Delta_i, \cfail, 2)$ & if $t_{\xmark, i} = \cmark$ & \Comment(``increase to complement'')\\
+						$dd_4(\cpass \cup \Delta_i, \cfail, max(n-1, 2))$ & if $t_{\cmark, i} = \cmark$ & \Comment(``increase to subset'')\\
+						$dd_4(\cpass, \cfail \setminus \Delta_i, max(n-1,2))$ & if $t_{\xmark, i} = \xmark$ & \Comment(``reduce to complement'')\\
+						$dd_4(\cpass, \cfail, min(2n, |\Delta|))$ & if $n < |\Delta|$ & \Comment(``increase granularity'')\\
+						(\cpass, \cfail) & otherwise & \Comment(``done'')
+					\end{tabular}}\right.$
+					(check the first cases for all $i \in \set{1, \dots, n}$)
+					\EndFunction
+				\end{algorithmic}
+		\end{minipage}
+	}
+	\caption{The Delta Debugging algorithm for finding two configurations with minimal difference, \dddiff.
+It finds $(\cfail, \cpass) = \dddiff(\C)$ such that $\emptyset \subseteq \cpass \subsetneq \cfail \subseteq \C$, $test(\cpass) = \cmark$, $test(\cfail) = \xmark$, and $\Delta = \cfail \setminus \cpass$ is 1-minimal.}
+	\label{fig:dddiff}
+\end{figure}
 
@@
 
-If a certain input is known to cause a program to fail execution Delta Debugging can be used to analyze which part of the program input is responsible for the failure.
+If a certain input is known to cause a program to fail execution then Delta Debugging can be used to analyse which part of the program input is responsible for the failure.
 The \ddp\ algorithm as defined above can be used without any modifications.
@@
 $\C$ is comprised of all tokens in \reffail.
-As depicted in \ref{fig:ddponfail} \ddp\ only takes 21 tests to find out that the C tokens "$+ 1.0$" are a minimal failure-inducing set of changes.
+As depicted in Figure \ref{fig:ddponfail} \ddp\ only takes 21 tests to find out that the C tokens "$+ 1.0$" are a minimal failure-inducing set of changes.
 \\
@@
 		\begin{minipage}{\textwidth}
-		\begin{centering}
+		\begin{centering}\texttt{
 					\begin{algorithmic}[1]
@@
 							\State $\}$
-			\end{algorithmic}
+					\end{algorithmic}}
 		\end{centering}
@@
 
-It can be of great help for the software developer to be able to automatically determine the specific parts of the program input that cause the failure.
+It can be of great help for the software developer to automatically determine the specific parts of the program input that cause the failure.
 The information gathered from the above method can be used by the developer to manually trace the responsible input data to the crash.
@@
 Delta Debugging can help trace the error from the root to the failure during program execution as well, though.
-Given a test framework that is able to access \hyperref[def:program_state]{program states} during execution, DD is used on the program states of two \hyperref[def:program_run]{program runs} \rpass\ and \rfail\ at significant locations to deduce a \textit{Cause-Effect-Chain (CEC)}.
+Given a debugging framework that is able to access \hyperref[def:program_state]{program states} during execution, DD is used on the program states of two \hyperref[def:program_run]{program runs} \rpass\ and \rfail\ at significant locations to deduce a \textit{Cause-Effect-Chain (CEC)}.
 
@@
 \definition \label{def:program_run} \defsub{Circumstances and program runs} Let $R$ be the set of all possible circumstance configurations.
-A run is defined as a configuration of circumstances, so every $r \in R$ is called a run.\\
+A (program) run is defined as a configuration of circumstances, so every $r \in R$ is called a run.\\
 
 For our intents and purposes it is sufficient to only consider the program input as the changeable circumstances as we assume the error lies within the program code and occurs independently of operating systems, hardware and other environmental factors.
+Thus, in our case a program run is equivalent to just the specific input that induces the program run.
 
@@
 We further assume knowledge of a passing program run \rpass\ that runs and terminates correctly.
-	\item We use Delta Debugging to compare program states of \rfail\ and \rpass\ at multiple significant locations \footnote{The number of comparisons vary depending on program size.
+	\item We use Delta Debugging repeatedly to compare program states of \rfail\ and \rpass\ at multiple significant locations \footnote{The number of comparisons vary depending on program size.
 It defines the granularity of the resulting Cause-Effect-Chain.
 The location of a comparison should give an overview of the state of the execution and should be reached by most program runs.
-The locations define the informative power of the Cause Effect Chain.}.
+The locations define the informative power of the Cause-Effect-Chain.}.
 Each time Delta Debugging is used the program state \sfail\ of \rfail\ is regarded as \td\ and the state \spass\ of \rpass\ is regarded as \yd.
 	\item A change between two program states is defined as either a different value of a variable or a variable that only exists in one program state.
-	\item The set $\C$ contains all such changes between \spass\ and \sfail\ for each comparison.
+	\item The set $\C$ contains all such changes between \spass\ and \sfail\ for each comparison, i.e. each individual execution of Delta Debugging.
 	\item The \textit{test} function models further program execution from a given program state.
@@
 
-The definition of the test function makes evident why a testing framework is needed.
+The definition of the test function makes evident why a debugging framework is needed.
 We need to be able to extract the program state \spass\ of \rpass\ and \sfail\ of \rfail\ at a certain point in order to compare them.
@@
 they are created by applying arbitrary subsets of $\C$ to \spass\ and would never occur in any run without outside interference.
-In that case unexpected behavior will occur and memory leaks are likely.
+In that case unexpected behaviour is inevitable and memory leaks are likely.
 The framework needs to make sure no harm is done to the machine or other software.
-Frameworks for testing and analyzing program execution with these capabilities luckily do exist already, though.
 
@@
 Applying GCC on $fail.c$ is regarded as \rfail\ and applying GCC to an empty C program as \rpass\ (compiling no code obviously terminates correctly).
-Andreas Zeller's experimental \textit{HOWCOME} algorithm\footnote{The \textit{HOWCOME} algorithm uses Delta Debugging as described above and incorporates technical issues like the extraction of program states and applying the test and analysis framework \cite{Zeller:2002:ICC:587051.587053}.}
-compares the runs at the function \textit{main} shortly after the start of the program, at the function $combine\_instructions$ in the middle of the program run and at the last hit of the function $if\_then\_else\_cond$ shortly before the failure.
-\textit{HOWCOME} yields a \hyperref[fig:failcec]{Cause-Effect-Chain} that includes the comparisons at those locations.
+Andreas Zeller's experimental \textit{HOWCOME} algorithm\footnote{The \textit{HOWCOME} algorithm uses Delta Debugging as described above and incorporates technical issues like the extraction of program states and applying the debugging framework \cite{Zeller:2002:ICC:587051.587053}.}
+compares the runs at the function \textit{main} shortly after the start of the program, at the function $combine\_instructions$ and at each hit of the function $if\_then\_else\_cond$.
+\textit{HOWCOME} yields a \hyperref[fig:failcec]{Cause-Effect-Chain} that includes the comparisons at those locations as shown in Figure \ref{fig:failcec}.
 
@@
 	}
-	\caption{The Cause-Effect-Chain that is deduced by \textit{HOWCOME} for \reffail\ \cite{Zeller:2002:ICC:587051.587053}.}
+	\caption{The Cause-Effect-Chain that is deduced by \textit{HOWCOME} for \reffail\ \cite{Zeller:2002:ICC:587051.587053}. Only the last hit of $if\_then\_else\_cond$ is shown.}
 	\label{fig:failcec}
@@
 The test driver will use the generated sequences to test the functionality of the UUT even under implausible circumstances.
-Non-randomized test drivers with test cases written by the programmer will not be able to cover all these cases.
+Non-randomized test drivers with test cases written by the programmer will not be able to cover all these cases \cite{Lei:2005:MRU:1104997.1105255}.
 
@@
 After generating random test cases it would be useful to minimize the failing test cases so they only contain the instructions that are relevant for inducing the failure.
-This is where Delta Debugging comes to play (\cite{Lei:2005:MRU:1104997.1105255}): By viewing a empty test case as the \yd\ state and the randomly generated failing test case as the \td\ state, we can apply a Delta Debugging algorithm to minimize the test case.
+This is where Delta Debugging comes to play (\cite{Lei:2005:MRU:1104997.1105255}): By viewing an empty test case as the \yd\ state and the randomly generated failing test case as the \td\ state, we can apply a Delta Debugging algorithm to minimize the test case.
 We define the abstract DD concepts in this context more concretely:
@@
 Both \hyperref[fig:ddp]{\ddp} and \hyperref[fig:dddiff]{\dddiff} can be used in this scenario.
-\ddp\ will return a minimal sequence of instruction that induces the same error as the original test case whereas \dddiff\ returns two test cases, one passing and one failing, with a minimal difference.
+\ddp\ will return a minimal sequence of instructions that induces the same error as the original test case whereas \dddiff\ returns two test cases, one passing and one failing, with a minimal difference.
 \dddiff\ will likely be more efficient (\cite{Zeller:2002:SIF:506201.506206})
@@
 At the same time static slicing is not as effective as DD minimization.
-A combination of both, that is first applying static slicing and then DD minimization, will retain the advantages of both methods, though: It has a lower runtime but is just as efficient as DD minimization \cite{Leitner:2007:EUT:1321631.1321698}.
+A combination of both, i.e.
+first applying static slicing and then DD minimization, will retain the advantages of both methods, though: It has a lower runtime but is just as efficient as DD minimization \cite{Leitner:2007:EUT:1321631.1321698}.
 
